### 2019-03-23 reports/thesis/chapters/5_inference.tex
Relocate inference figures

The aim of this project was to obtain an x-ray image of a 3D printed sample and compare it with a simulation of that scan using software called \texttt{aRTist} \citep{bellon2007artist, jaenisch2008artist, bellon2012radiographic}. \texttt{aRTist} can simulate x-ray acquisitions of the 3D printed sample given the specifications of the scan, such as the x-ray source, the x-ray detector and the computer aided design (CAD) of the 3D printed sample \citep{bellon2011simulation, deresch2012simulating}. Users of \texttt{aRTist} can align the simulated x-ray image to the real x-ray image using numerical methods, however this is outside the scope of this thesis.

Disagreement between the scan and \texttt{aRTist} can be found by simply subtracting one image from the other, any values too big in magnitude can be considered as a defect. However in the previous chapters, it was found that x-ray photons behave randomly and differences in the comparison can be due to chance. Thus the comparisons should be done under the face of uncertainty.

A pixel by pixel inference was proposed to do defect detection. A statistic $Z_{x,y}$ for the $(x,y)$ positioned pixel was calculated for all pixels. This statistic is
\begin{equation}
  Z_{x,y} = 
  \dfrac{
    \text{scan}_{x,y} - \texttt{aRTist}_{x,y}
  }
  {
    \sqrt{\widehat{\variance}\left[\text{scan}_{x,y}\right]}
  }
\end{equation}
where $\widehat{\variance}\left[\text{scan}_{x,y}\right]$ is the estimated grey value variance of pixel $(x,y)$ in the scan. Under mild assumptions, it was shown in the previous chapters that the grey values in the scan were Normal. Thus by treating the simulated image from \texttt{aRTist} as known and the ground truth, then the randomness of the $Z_{x,y}$ statistics, assuming there are no defects, can be approximately quantified
\begin{equation}
Z_{x,y}\sim \normal(0,1) \ .
\end{equation}
As with usual statistical convention, upper case $Z$ denote a random variable. Whereas lower case $z$ denote a realisation or an observation of that random variable.

The variance model, $\widehat{\variance}\left[\text{scan}_{x,y}\right]$, can be calibrated, or trained, by holding out a number replicated x-ray images of a 3D printed sample. These replicated x-ray images were manually segmented. This provided variance-mean data and was used to train the variance model, such as a Gamma distributed GLM. The method on doing so was described in the previous chapter and it was found a linear relationship between the variance and the mean was a good model. The variance was then predicted using the grey value in the \texttt{aRTist} simulation as the predictor variable.

This method for inference was applied to the dataset \texttt{Sep16 120deg}. Here a cuboid with voids were purposefully manufactured to see if they can be detected, the CAD shown in Figure \ref{fig:inference_testObject}. The 20 x-ray images were spilt into 2. 19 images were used to train the variance-mean model. One image, called the test image, was used to compare with \texttt{aRTist}. In the \texttt{aRTist} simulation, the software did a simulation as if the voids were not there. The test image and the \texttt{aRTist} simulation is shown in Figures \ref{fig:inference_inferenceIntro_inference_scan} and \ref{fig:inference_inferenceIntro_inference_artist}.

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{../figures/inference/TestObject.png}
  \caption{CAD of a 3D cuboid with voids}
  \label{fig:inference_testObject}
\end{figure}

\begin{figure}
  \centering
  \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[width=\textwidth]{../figures/inference/inferenceIntro_scan.eps}
      \caption{X-ray image}
      \label{fig:inference_inferenceIntro_inference_scan}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[width=\textwidth]{../figures/inference/inferenceIntro_artist.eps}
      \caption{\texttt{aRTist} simulation}
      \label{fig:inference_inferenceIntro_inference_artist}
    \end{subfigure}
  }
  \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[width=\textwidth]{../figures/inference/inferenceIntro_logp.eps}
      \caption{$-\log p\text{-values}$}
      \label{fig:inference_inferenceIntro_inference_pvalue}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[width=\textwidth]{../figures/inference/inferenceIntro_positivePixels.eps}
      \caption{Positive pixels}
      \label{fig:inference_inferenceIntro_inference_positivePixels}
    \end{subfigure}
  }
  \caption{An x-ray image of a 3D printed cuboid, from the \texttt{Sep16 120deg} dataset, was compared to the \texttt{aRTist} simulation to detect the purposefully manufactured voids. The $p$-values obtained were used for hypothesis testing using the \cite{benjamini1995controlling} procedure, controlling the FDR at the 5\% level. Pixels detected as positive are shown in red in d).}
  \label{fig:inference_inferenceIntro_inference}
\end{figure}

For each pixel, a $z$ statistic was calculated. A $z$ statistic too large in magnitude will classify that pixel as a positive result. In two-tailed hypothesis testing framework \citep{pearson1900on, neyman1933on, fisher1970statistical}, the $p$-value can be used to represent $z$ statistic in a different way
\begin{equation}
  p_{x,y} = 2(1-\Phi(|z_{x,y}|))
\end{equation}
which can takes values $0\leqslant p_{x,y} \leqslant 1$. A $p$-value too small is considered a positive result. The $p$-values are shown in Figure \ref{fig:inference_inferenceIntro_inference_pvalue}.

The resulting $p$-values are concerning. This is because the $p$-values are not very smooth on the surfaces of the sample. It should be expected that small $p$-values are in areas of the defects. Pixels are tested positive, or considered to be evidence of a defect, when $|z_{x,y}|>\inputNumber{../figures/inference/inferenceIntro_critical.txt}$. This value was chosen by controlling the false discovery rate at 5\% using the \cite{benjamini1995controlling} (BH) procedure. The positive pixels are shown in Figure \ref{fig:inference_inferenceIntro_inference_positivePixels}.

This proposed method for defect detection failed because too many false positives were detected. These false positives appear to have some structure, for example clustering in the corners or on surfaces. In addition, false negatives were detected because not all of the defects were detected.

Model misspecification appeared to be the main source of error. The $z$ statistics and $p$-values were inspected in Figure \ref{fig:inference_inferenceIntro_histogram}. It can be seen that the majority of the $z$ statistics does not look Normal. Also the majority of the $p$-values does not look uniformly distributed. This seems to suggest that the assumption of $Z_{x,y}\sim \normal(0,1)$ is incorrect.

\begin{figure}
  \centering
  \centerline{
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/inferenceIntro_histogram.eps}
    \caption{Histogram of $z$}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/inferenceIntro_pValue.eps}
    \caption{$p$-values}
    \label{fig:inference_inferenceIntro_pvalue}
  \end{subfigure}
}
  \caption{a) The histogram of the $z$ statistics is compared with the standard Normal distribution. The standard Normal distribution is known as the null distribution in this scenario. b) The $p$-values are ordered and plotted. The critical region corresponds to controlling the false discovery rate at 5\% using the \cite{benjamini1995controlling} procedure.}
  \label{fig:inference_inferenceIntro_histogram}
\end{figure}

This chapter recaps hypothesis testing, for a single test and then for multiple tests, treating each pixel as a test. Assumptions used on the hypothesis testing can be relaxed by using the empirical null \citep{efron2004large}, which is reviewed here. The empirical null is then extended to an image filter, called the empirical null filter. This filter adjust each $z$ statistic according to its neighbours, ironing out false positive results. Simulations and results are shown towards the end of the chapter.

\section{Literature Review}

Consider a $z$ statistic from a pixel. If there are no defects, then the statistic is null and has the null distribution $Z|H_0\sim\normal(0,1)$. This can be described in another way by specifying the random variable as $Z\sim\normal(\mu,1)$ and the null hypothesis as $H_0:\mu=0$. A hypothesis test can judge how much $Z$ deviates from 0 by defining the alternative hypothesis $H_1:\mu\neq 0$. A statistic which does not have the null distribution is known as non-null.

In two-tailed hypothesis testing \cite{pearson1900on, neyman1933on, fisher1970statistical}, the $p$-value can be compared with the user defined significance level $\alpha$, also known as the size of the test in this specific example. A positive result is declared when $p<\alpha$, else it is a negative result. Correct and incorrect testing of the statistics can occur. For example testing a null statistic as positive is known as a false positive. It can be shown that $\alpha$ controls the false positive rate. $\alpha=5.0\%$ is a typical choice and was used throughout this thesis. Other fields use the $\sigma$ notation, for example $2\sigma$ correspond to about the $\alpha = 4.6\%$ significance level; this notation will not be used here.

Multiple hypothesis testing occurs when there are more than one hypothesis to test. For example $n$ pixels can be tested where the $z$ statistics are $Z_1, Z_2, \cdots, Z_n$ and $Z_i\sim\normal(\mu_i,1)$ for $i=1,2,\cdots,n$. The null hypotheses are $H_{0,i}:\mu_i=0$ and are tested against the alternative $H_{1,i}:\mu_i\neq 0$ for $i=1,2,\cdots,n$. Let the corresponding $p$-values be $p_1, p_2, \cdots, p_n$.

The uncorrected test classify any $p_i<\alpha$ as positive. This is flawed because the possibility of obtaining at least one false positive increases as the number of tests increase \citep{shaffer1995multiple}. This method does control the per comparison error rate (PCER) \citep{benjamini1995controlling}. PCER is the proportion of false positives out of all tests. The number of true/false positive/negatives obtained are defined in Table \ref{table:inference_randomvariables}. Using the notation, PCER is defined as
\begin{equation}
  \text{PCER}=
  \dfrac{1}{n}
  \expectation[V]
  \ .
\end{equation}
It can be seen that if the uncorrected test controls the false positive rate such that $\expectation[V]/n_0 = \alpha$, then it controls the PCER such that
\begin{equation}
  \text{PCER}\leqslant\alpha \ .
\end{equation}

\begin{table}
  \centering
  \begin{tabular}{c|c|c|c}
    &Negative&Positive&Total\\\hline
    Null & $U$ & $V$ & $n_0$\\
    Non-null & $T$ & $S$ & $n-n_0$\\\hline
    &$n-R$&$R$&$n$
  \end{tabular}
  \caption{Random variable definitions for true/false positives/negatives made in multiple hypothesis testing}
  \label{table:inference_randomvariables}
\end{table}

The Bonferroni correction \citep{shaffer1995multiple, bland1995multiple, perneger1998what} controls the family-wise error rate (FWER) \citep{shaffer1995multiple} where
\begin{equation}
  \text{FWER} = 1-\prob(V=0) \ .
  \label{eq:inference_fwer}
\end{equation}
This is done by adjusting the size of the test to be $\alpha/n$. By using the adjusted size, then $\text{FWER}=1-\left[(1-\alpha/n)^n\right]$. Using the approximation $(1-\alpha/n)^n\approx 1-\alpha$ then $\text{FWER}\approx \alpha$. This shows that the Bonferroni correction controls the family-wise error rate such that
\begin{equation}
  \text{FWER} \leqslant \alpha \ .
\end{equation}
However in practice the Bonferroni correction is not very powerful \citep{perneger1998what}, meaning it does give too many false negatives. This is because the correction traded too many false positives for false negatives.

The \cite{benjamini1995controlling} (BH) procedure controls the false discovery rate (FDR) \citep{benjamini2010discovering} rather than the PCER or FWER. The FDR is the proportion of false positives out of all positive results, that is
\begin{equation}
  \text{FDR} = \expectation\left[\dfrac{V}{R}\right]
  \ .
\end{equation}
It is defined that $V/R=0$ when $R=0$.

The BH procedure adjust the size of the test between the Bonferroni correction and the uncorrected test. It adapts to the data and choose different sizes for different data. The procedure is as follows, the $p$-values are ordered such that $p_{(1)}\leqslant p_{(2)}\leqslant \cdots \leqslant p_{(n)}$. Suppose a size $\alpha$ is provided beforehand. The size is then adjusted to
\begin{equation}
  \alpha_{\text{BH}} = \frac{\alpha k}{n}
\end{equation}
where
\begin{equation}
  k\text{ is the largest }i\text{ for which }p_{(i)}\leqslant\frac{i}{n}\alpha
  \ .
\end{equation}
This test the statistics with $p$-values $p_{(1)},p_{(2)},\cdots,p_{(k)}$ as positive. For the case where $p_{(1)}>\alpha/n$ then $k=1$ so that Bonferroni correction is used when there are no positive results, this is only for illustration purposes. For example in Figure \ref{fig:inference_inferenceIntro_pvalue} the $p$-values were plotted against their order so that the critical boundary was shown by a linear curve with gradient $\alpha/n$.

The BH procedure comes from the fact that if all the statistics are null and independent, then the $p$-values are uniformly distributed \citep{simes1986improved}. However it can be shown that the BH procedure works for many scenarios of dependencies \citep{benjamini2001control}. It can be shown that the BH procedure controls the FDR such that
\begin{equation}
  \text{FDR}\leqslant\alpha
\end{equation}
\citep{benjamini1995controlling}.

In summary, the uncorrected, Bonferroni and BH correction controls for different error rates according to the threshold $\alpha$, summarised in Table \ref{table:inference_corrections}.

\begin{table}
    \centering
    \begin{tabular}{c|c}
        Correction&Controls for\\\hline
        No correction&$\text{PCER}\leqslant\alpha$\\
        Bonferroni&$\text{FWER}\leqslant\alpha$\\
        BH&$\text{FDR}\leqslant\alpha$
    \end{tabular}
    \caption{Different types of corrections for multiple hypothesis testing are listed here, along with what they control for.}
    \label{table:inference_corrections}
\end{table}

\begin{figure}
  \centering
    \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[width=\textwidth]{../figures/inference/inferenceSubsample1_zImage.eps}
      \caption{$z$ statistics}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[width=\textwidth]{../figures/inference/inferenceSubsample1_histogram.eps}
      \caption{Histogram of $z$ in sub-image}
    \end{subfigure}
    }
    \caption{The resulting $z$ image is shown in a). A $200\times 200$ sub-image was taken shown by the dashed lines. A histogram of the $z$ statistics in the sub-image is shown in b). The critical region was obtained by using the BH procedure at the 5\% FDR level.}
    \label{fig:inference_inferenceSubsample1_histogram}
\end{figure}

Consider a small example where a $200\times200$ section of the $z$ image was investigated, as shown Figure \ref{fig:inference_inferenceSubsample1_histogram}. The distribution of the $z$ statistics looked Normal but not centred at zero, as a result using the BH procedure obtained a critical region of $|Z|>\input{../figures/inference/inferenceSubsample1_criticalBoundary.txt}$. Almost half of the pixels were tested positive. It is questionable whether such a number of positive results is sensible, in particular in an area where defects were not expected.

This problem commonly occur in large scale multiple hypothesis testing \citep{efron2004large} such as in microarrays \citep{hedenfalk2001gene, efron2002empirical, efron2003robbins}. It appears that the null distribution was misspecified as $Z_i|H_{0,i}\sim\normal(0,1)$ when in reality they are distributed as $Z_i|H_{0,i}\sim\normal(\mu_0,\sigma_0^2)$, where $\mu_0$ and $\sigma_0$ are the null mean and null standard deviation respectively.

In practice $\mu_0$ and $\sigma_0$ are unknown. The empirical null \citep{efron2004large} replaces the parameters of the null distribution with its estimate. Suppose $\widehat{\mu}_0$ and $\widehat{\sigma}_0$ are the estimated null mean and null standard deviation respectively. The statistics are specified by $Z_i\sim\normal(\mu_i, \widehat{\sigma}_0^2)$ so that the following null hypotheses $H_{0,i}:\mu_i=\widehat{\mu}_0$ are tested against $H_{1,i}:\mu_i\neq\widehat{\mu}_0$ for $i=1,2,\cdots,n$.

In \cite{efron2004large}, a few assumptions were made in order to obtain the estimators $\widehat{\mu}_0$ and $\widehat{\sigma}_0$. Let the test statistic $Z$ have the p.d.f.
\begin{equation}
  p_Z(z) =
  \pi_0 p_{Z|H_0}(z) + \pi_1 p_{Z|H_1}(z)
\end{equation}
where $0\leqslant\pi_0\leqslant 1$ and  $\pi_1 = 1-\pi_0$. In addition, the null distribution is
\begin{equation}
  p_{Z|H_0}(z) = 
  \dfrac{1}{\sqrt{2\pi}\sigma_0}
  \exp\left[
    -\dfrac{1}{2}
    \left(
      \dfrac{z-{\mu}_0}{\sigma_0}
    \right)^2
  \right]
\end{equation}
as it was assumed to be Normal. The non-null distribution $p_{Z|H_1}(z)$ does not need to be specified. Assume that the majority of the data are null and non-null test statistics are rare, say $\pi_0>0.9$ \citep{efron2004large}. Then around the mode, the probability density function would be dominated by the null distribution. This implies that
\begin{equation}
  p_{Z}(z) \approx \pi_0 p_{Z|H_0}(z)
\end{equation}
for values of $z$ around the mode. Finding the mode for $p_{Z}(z)$ and $p_{Z|H_0}(z)$ should yield the same solution. This justify the use of the mode for the empirical null mean \citep{efron2004large}
\begin{equation}
  \widehat{\mu}_0 = \argmax\widehat{p}_Z(z)
\end{equation}
where $\widehat{p}_Z(z)$ is the density estimation of $p_Z(z)$, for example it could be a smoothing spline fitted onto the histogram \citep{efron2004large}.

The null standard deviation was estimated from the log density \citep{efron2004large}. For values of $z$ at and around the mode
\begin{equation}
  \ln p_{Z}(z) = 
  \ln\left[
    \dfrac{\pi_0}{\sqrt{2\pi}{\sigma}_0}
  \right]
  -\dfrac{1}{2}
  \left(
    \dfrac{z-{\mu}_0}{{\sigma}_0}
  \right)^2
  \ .
\end{equation}
Taking derivatives
\begin{equation}
  \dfrac{\partial}{\partial z} \ln p_{Z}(z) =
  -\left(
    \dfrac{
      z-{\mu}_0
    }
    {
      {\sigma}_0^2
    }
  \right)
\end{equation}
\begin{equation}
  \dfrac{\partial^2}{\partial z^2} \ln p_{Z}(z) =
  -\dfrac{
    1
  }
  {
    {\sigma}_0^2
  }
\end{equation}
which motivates the estimator \citep{efron2004large}
\begin{equation}
  \widehat{\sigma}_0 = \left[
    \left.
      -\dfrac{\partial^2}{\partial z^2}\ln\widehat{p}(z)
    \right|_{z=\widehat{\mu}_0}
  \right]^{-1/2}
  \ .
\end{equation}
The evaluation of $z=\widehat{\mu}_0$ is used because at the mode, it was assumed that the null test statistics would dominate and non-null test statistics would not contribute much to the density estimate.

This method does not need estimations of $\pi_0$ which makes it quite useful. The estimations of $\pi_0$ is discussed in literature such as \cite{benjamini2000adaptive, pounds2003estimating, storey2003statistical, pounds2004improving, langaas2005estimating, durnez2014posthoc}. \cite{efron2004large} motives the use of the empirical null by investigating various types of false discovery rates \citep{storey2002direct, storey2003positive, efron2002empirical, efron2007size} which will not be discussed here.

The empirical null is used widely, including in neuroimaging \citep{schwartzman2008false, schwartzman2009empirical} and has been extended to include non-Normal null distributions \citep{schwartzman2008false, schwartzman2008empirical}. There exist various methods for the empicial null. \cite{schwartzman2008empirical} uses Poisson regression on the histogram counts in regions where the null statistics dominate. The empirical characteristic function was used in \cite{jin2007estimating}.

\section{Empirical Null}

Return back to the sub-image example in Figure \ref{fig:inference_inferenceSubsample1_histogram}. The histogram was smoothed using a kernel density estimate \citep{parzen1962on, friedman2001elements} as shown in Figure \ref{fig:inference_inferenceSubsample1_densityEstimate}. The kernel density estimate \citep{parzen1962on} is
\begin{equation}
  \widehat{p}_Z(z)=
  \frac{1}{nh}
  \sum_{i=1}^n\phi\left(
    \dfrac{z_i-z}{h}
  \right)
  \label{eq:inference_kernel_density_estimate}
\end{equation}
where $h$ is the bandwidth. The bandwidth was chosen such that
\begin{equation}
  h = 0.9n^{-1/5}\times\text{min}\left(s_z,\text{IQR}_z/1.34\right) + 0.16
  \label{eq:inference_ourruleofthumb}
\end{equation}
where $s_z$ and $\text{iqr}_z$ are the standard deviation and interquartile range.

\begin{figure}
  \centering
  \includegraphics[width=0.49\textwidth]{../figures/inference/inferenceSubsample1_densityEstimate.eps}
  \caption{The density of the $z$ statistics was estimated using a kernel density estimate. The dot dash lines shows the empirical null distribution multiplied by some constant, this is to insulate the two densities have the same curvature at the mode.}
  \label{fig:inference_inferenceSubsample1_densityEstimate}
\end{figure}

The null parameters were estimated. The null mean was estimated by numerically finding the mode of the density estimate. The null standard deviation was estimated using the curvature of the density estimate at the mode. This resulted in the empirical null density to have the same curvature as the density estimate at the mode, this is illustrated in Figure \ref{fig:inference_inferenceSubsample1_densityEstimate}. In this particular example, it was found that $\widehat{\mu}_0=\input{../figures/inference/inferenceSubsample1_nullMean.txt}$ and $\widehat{\sigma}_0=\input{../figures/inference/inferenceSubsample1_nullStd.txt}$ to 2 decimal places.

The use of the kernel density estimate has its advantages. The only tuning parameter is the bandwidth $h$ and the density estimate is simple enough to do calculus on it. Speed may be an issue as an evaluation of the density estimate requires the sum over the $n$ $z$ statistics.

The $z$ statistics were normalised to $T_1,T_2,\cdots,T_n$ by using
\begin{equation}
  T_i = \dfrac{
    Z_i - \widehat{\mu}_0
  }
  {
    \widehat{\sigma}_0
  }
  \label{eq:inference_empiricalNullNormalise}
\end{equation}
for $i=1,2,\cdots,n$. The normalised $p$-values were obtained by using
\begin{equation}
  p_i = 2(1-\Phi(|t_i|))
\end{equation}
and then were used in the BH procedure. The critical region was found to be $|t|\geqslant\inputNumber{../figures/inference/inferenceSubsample1_nullNormalisedCritical.txt}$. In terms of the original units, the critical region was $z \leqslant \inputNumber{../figures/inference/inferenceSubsample1_nullCritical1.txt}$ and $\inputNumber{../figures/inference/inferenceSubsample1_nullCritical2.txt}\leqslant z$, all to 2 decimal places. The critical region and the normalised $p$-values are shown in Figure \ref{fig:inference_inferenceSubsample1_nullResults}.

\begin{figure}
  \centering
    \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/inferenceSubsample1_nullHistogram.eps}
        \caption{Histogram of $z$}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/inferenceSubsample1_nullPValues.eps}
        \caption{$p$ values}
    \end{subfigure}
    }
    \caption{The critical regions and $p$-values were adjusted using the empirical null.}
    \label{fig:inference_inferenceSubsample1_nullResults}
\end{figure}

No positive results were found and the $p$-values were sensible as they looked uniformly distributed. This demonstrated that the empirical null adjusted the parameters of the null distribution to fit the majority of the data in order to make sensible inference.

\begin{figure}
  \centering
    \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/inferenceSubsample2_zImage.eps}
        \caption{$z$ statistics}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/inferenceSubsample2_subimagePositive.eps}
        \caption{$z$ from the sub-image}
    \end{subfigure}
    }
    \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/inferenceSubsample2_nullHistogram.eps}
        \caption{Histogram of $z$}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/inferenceSubsample2_nullPValues.eps}
        \caption{$p$ values}
    \end{subfigure}
    }
    \caption{A $200\times 200$ sub-image containing a defect is shown in a). The empirical null was used to test pixels, positive pixels are shown in red in b). The histogram of the $z$ statistics are shown in c) along with the critical region. The normalised $p$-values are shown in d). The critical regions were adjusted using the empirical null and obtained using the BH procedure, controlling the FDR at 5\%.}
    \label{fig:inference_inferenceSubsample2}
\end{figure}

Another example is shown in Figure \ref{fig:inference_inferenceSubsample2} where a $200\times 200$ sub-image containing a defect was used. It can be seen that the null distribution was not centred at zero, however the empirical null took it into account. The estimation of the empirical null parameters were robust as it only depended on the density estimate at the mode only, it should not be affected by non-null test statistics.

The BH procedure was conducted using the normalised $p$-values. $\inputNumber{../figures/inference/inferenceSubsample2_nPositive.txt}$ pixels were tested positive. The majority of the positive pixels were found to be clustered together, which highlighted the defect. Not all of the defects were tested positive, but a good number of them are. This should be enough to raise suspicion in that particular area. Only a few pixels were falsely tested as positive, however they are typically isolated single pixels. Isolated positive pixels should be discarded as they are more than likely to be tested positive by random chance.

This section describes the empirical null implementation in detail, in particular the numerical and computational aspects. The empirical properties of the sampling distribution of the null parameter estimators were studied.

\subsection{Mode Finding}

The mode was found by solving $\widehat{\mu}_0 = \argmax\widehat{p}_Z(z)$. This cannot be solved in closed form easily and was done numerically. The  Newton-Raphson method can be used to solve $\hat{p}_Z'(z) = 0$ for $z$. Instead the log space was used, solving
\begin{equation}
  \dfrac{
    \partial
  }
  {
    \partial z
  }
  \ln\widehat{p}_Z(z)
  = 0
\end{equation}
for $z$. This at least find stationary points.

The Newton-Raphson method is an iterative algorithm and requires an initial value $z^{(0)}$. The iterative step is
\begin{equation}
  z^{(r+1)} =
  z^{(r)}
  -\dfrac{
    \left.
      \dfrac{
        \partial
      }
      {
        \partial z
      }
      \ln\widehat{p}_Z(z)
    \right|_{z = z^{(r)}}
  }
  {
    \left.
      \dfrac{
        \partial^2
      }
      {
        \partial z^2
      }
      \ln\widehat{p}_Z(z)
    \right|_{z = z^{(r)}}
  } 
\end{equation}
for $r=0,1,2,3,\cdots$ until some convergence condition is met. The derivatives for the log density was obtained via the following. Recall that the kernel density estimate is $\widehat{p}_Z(z)=
  \frac{1}{nh}
  \sum_{i=1}^n\phi\left(
    \frac{z_i-z}{h}
  \right) $
so that the log density is
\begin{equation}
  \ln\widehat{p}_Z(z)=
  \ln\left(
    \dfrac{1}{nh}
  \right)
  +
  \ln\left[
    \sum_{i=1}^n
    \phi\left(
      \dfrac{
        z_i - z
      }
      {
        h
      }
    \right)
  \right]
  \ .
\end{equation}
Taking the first derivative
\begin{equation*}
  \dfrac{
    \partial
  }
  {
    \partial z
  }
  \ln\widehat{p}_Z(z)
  =
  \dfrac{
    1
  }
  {
    \sum_{i=1}^n
    \phi\left(
      \dfrac{
        z_i - z
      }
      {
        h
      }
    \right)
  }
  \times
  \sum_{i=1}^n
  \phi'\left(
    \dfrac{
      z_i - z
    }
    {
      h
    }
  \right)
  \left(
    -\dfrac{
      1
    }
    {
      h
    }
  \right)
  \ .
\end{equation*}
Using the fact that $\phi(z)=(2\pi)^{-1/2}\exp(-z^2/2)$, then $\phi'(z)=-z\phi(z)$. This is used to simplify the equation to be
\begin{equation}
  \dfrac{
    \partial
  }
  {
    \partial z
  }
  \ln\widehat{p}_Z(z)
  =
  \dfrac{
    \sum_{i=1}^n
    \phi\left(
      \dfrac{
        z_i - z
      }
      {
        h
      }
    \right)
    \left(
      \dfrac{
        z_i - z
      }
      {
        h
      }
    \right)
  }
  {
    h
    \sum_{i=1}^n
    \phi\left(
      \dfrac{
        z_i - z
      }
      {
        h
      }
    \right)
  }
  \ .
\end{equation}
Taking the derivative again
\begin{multline*}
  \dfrac{
    \partial^2
  }
  {
    \partial z^2
  }
  \ln\widehat{p}_Z(z)
  =
  \left[
    h\sum_{i=1}^n
    \phi\left(
      \dfrac{
        z_i-z
      }
      {
        h
      }
    \right)
  \right]^{-2}
  \times
  \left\{
    h\left[
      \sum_{i=1}^n
      \phi\left(
        \dfrac{
          z_i-z
        }
        {
          h
        }
      \right)
    \right]
  \right.
  \\
  \left.
    \times
    \sum_{i=1}^n\left[
      \phi'\left(
        \dfrac{
          z_i-z
        }
        {
          h
        }
      \right)
      \left(
        -\dfrac{
          1
        }
        {
          h
        }
      \right)
      \left(
        \dfrac{
          z_i-z
        }
        {
          h
        }
      \right)
      +\phi\left(
        \dfrac{
          z_i-z
        }
        {
          h
        }
      \right)
      \left(
        -\dfrac{
          1
        }
        {
          h
        }
      \right)
    \right]
  \right.
  \\
  \left.
    -
    \left[
      \sum_{i=1}^n
      \phi\left(
        \dfrac{
          z_i-z
        }
        {
          h
        }
      \right)
      \left(
        \dfrac{
          z_i-z
        }
        {
          h
        }
      \right)
    \right]
    \left[
      h\sum_{i=1}^n
      \phi'\left(
        \dfrac{
          z_i-z
        }
        {
          h
        }
      \right)
      \left(
        -\dfrac{
          1
        }
        {
          h
        }
      \right)
    \right]
  \right\}
  \ .
\end{multline*}
Using the fact that $\phi'(z)=-z\phi(z)$, then it is simplified to
\begin{multline}
  \dfrac{
    \partial^2
  }
  {
    \partial z^2
  }
  \ln\widehat{p}_Z(z)
  =
  \left[
    h\sum_{i=1}^n
    \phi\left(
      \dfrac{
        z_i-z
      }
      {
        h
      }
    \right)
  \right]^{-2}
  \times
  \left\{
    \left[
      \sum_{i=1}^n
      \phi\left(
        \dfrac{
          z_i-z
        }
        {
          h
        }
      \right)
    \right]
  \right.
  \\
  \left.
    \times
    \left[
      \sum_{i=1}^n
      \phi\left(
        \dfrac{
          z_i-z
        }
        {
          h
        }
      \right)
      \left(
        \left(
          \dfrac{
            z_i-z
          }
          {
            h
          }
        \right)^2
        -1
      \right)
    \right]
    -
    \left[
      \sum_{i=1}^n
      \phi\left(
        \dfrac{
          z_i-z
        }
        {
          h
        }
      \right)
    \right.
  \right.
  \\
  \left.
    \left.
      \left(
        \dfrac{
          z_i-z
        }
        {
          h
        }
      \right)
    \right]^2
  \right\}
  \ .
\end{multline}

A convergence criteria was met when either 10 update steps were taken or when
\begin{equation}
  \log\left[\left|
    \left.
    \dfrac{
      \partial
    }
    {
      \partial z
    }
  \ln\widehat{p}_Z(z)
  \right|_{z=\widehat{\mu}_0}
  \right|\right]
  <-5
\end{equation}
at the current step. This was chosen arbitrary to speed up the algorithm without losing too much accuracy. At the end of the algorithm, for a successful convergence it was required, in addition, that
\begin{equation}
  \left.
    \dfrac{
      \partial^2
    }
    {
      \partial z^2
    }
    \ln\widehat{p}_Z(z)
  \right|_{z=\widehat{\mu}_0}
  < 0 \ .
\end{equation}
Following a successful convergence, the estimator $\widehat{\sigma}_0$ was calculated straight away.

The algorithm does depend on the initial value so using different initial values were used. Further initial values were generated by sampling from $\normal(z^{(0)}, s_z^2)$. This was done multiple times until 3 valid solutions were obtained. The best solution, the one with the largest $\ln\widehat{p}_Z\left(\widehat{\mu}_0\right)$, out of all the different initial values was used as the final answer. The Newton-Raphson method can fail if it converges to a stationary point other than a global maximum, for example a local maximum or a point of inflection.

\subsection{Bandwidth Tuning}

The bandwidth $h$ controls how smooth the kernel density estimator is, higher values producing smoother curves \citep{friedman2001elements}. Cross validation methods to select $h$ do exist \citep{bowman1984alternative, sheather2004density} but they can be computationally expensive. Rules of thumb \citep{silverman1986density, sheather2004density} can be used instead and are typically of the form
\begin{equation}
  h = bn^{-1/5}\times\text{min}\left(s_z,\text{IQR}_z/1.34\right)
\end{equation}
where $b=0.9$ \citep{silverman1986density}. This rule of thumb was developed with consideration of bimodal Normal distributions \citep{silverman1986density}. Other options include $b=1.06$ and $b=1.144$ to produce smoother curves \citep{silverman1986density, sheather2004density}.

The kernel density estimator was only used to obtain values of $\widehat{\mu}_0$ and $\widehat{\sigma}_0$. Thus a good bandwidth, in the context of the empirical null, is one which has good properties of $\widehat{\mu}_0$ and $\widehat{\sigma}_0$ rather than the density estimate. Exact properties of these estimators based on the kernel density estimators can be rather complicated. For example numerous approximations were needed to show that $\widehat{\sigma}_0^2$ is an unbiased estimator of $\sigma_0^2$ to the first order, where higher order terms are functions of $(\widehat{\mu}_0 - \mu_0)^2$ and $h$. This is shown in Appendix \ref{chapter:appendix_expectationNullStdEstimator}.

An experiment was conducted to investigate how the estimators, $\widehat{\mu}_0$ and $\widehat{\sigma}_0$, varied with $h$ and $n$. This was done on a simulated dataset of $n$ standard Normal random variables. For a given $h$ and $n$, 100 values of $\widehat{\mu}_0$ and $\widehat{\sigma}_0$ were obtained by repeating the simulation of the datasets. The median squared error was obtained by taking the median over all squared errors from the 100 estimates. Figure \ref{fig:inference_BandwidthSelection_error} shows the resulting median squared errors.

$\widehat{\mu}_0$ has low median squared error for large bandwidths. For low $n$, large bandwidths are particularly important because smoother curves prevents any false bimodal features appearing, making it easier to find the mode.

A smooth valley can be seen for the median squared error for $\widehat{\sigma}_0$. It appears for a given $n$, there exist a bandwidth which minimise the median squared error. However the rules of thumb does not optimise for $\widehat{\sigma}_0$ and undershoots it. Figure \ref{fig:inference_BandwidthSelection_error3} shows the median of the estimates of $\sigma_0$. The true value in this simulation is $\sigma_0=1$ which is represented as a horizontal plane. Bandwidths too small underestimate $\sigma_0$, bandwidths too big over estimates $\sigma_0$.

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[width=\textwidth]{../figures/inference/BandwidthSelection_error1.eps}
      \caption{$\mu_0$}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[width=\textwidth]{../figures/inference/BandwidthSelection_error2.eps}
      \caption{$\sigma_0$}
  \end{subfigure}
  \caption{Median squared error of the estimators of the empirical null parameters from 100 repeats of $n$ simulated $\normal(0,1)$ data. Lines represent the rule of thumb for various values of $b$.}
  \label{fig:inference_BandwidthSelection_error}
\end{figure}

\begin{figure}
  \centering
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/BandwidthSelection_error3.eps}
        \caption{Heatmap}
    \end{subfigure}
    \caption{Median value of $\widehat{\sigma}_0$ from 100 repeats of $n$ simulated $\normal(0,1)$ data. The true value is $\sigma_0=1$ which is represented by a horizontal plane.}
    \label{fig:inference_BandwidthSelection_error3}
\end{figure}

Overestimates and underestimates are both dangerous and it can affect the false positive and false negative rate. This is because $\widehat{\sigma}_0$ was used to rescale the test statistics accordingly. Recall that the normalised test statistics are $t_i = \dfrac{z_i - \widehat{\mu}_0}{\widehat{\sigma}_0}$. Low values of $\widehat{\sigma}_0$ produce large values in magnitude of $t_i$, potentially testing more of these test statistics as positive. High values of $\widehat{\sigma}_0$ does the opposite. Large number of positives can led to more false positives and vice versa.

The optimal bandwidth for a given $n$ was found numerically. Figure \ref{fig:inference_BandwidthSelection2_plot} shows the log squared error against the bandwidth for various $n$. A smoothing spline \citep{friedman2001elements} was fitted and this was optimised to find the bandwidth which minimises the log squared error for a given $n$. \texttt{MATLAB}'s \texttt{fit(x,y,'smoothingspline')} function was used to fit the smoothing spline which has inbuilt cross validation procedures. The optimisation was done using \texttt{MATLAB}'s gradient free \texttt{fminsearch} function, using $h=0.9n^{-1/5}$ as the initial value.

\begin{figure}
  \centering
    \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/BandwidthSelection2_plot1.eps}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/BandwidthSelection2_plot51.eps}
    \end{subfigure}
    }
    \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/BandwidthSelection2_plot100.eps}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/BandwidthSelection2_plot150.eps}
    \end{subfigure}
    }
    \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/BandwidthSelection2_plot199.eps}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/BandwidthSelection2_plot249.eps}
    \end{subfigure}
    }
    \caption{A smoothing spline was fitted on the log squared error for estimating $\sigma_0$ against the bandwidth. The estimator $\widehat{\sigma}_0$ uses the kernel density estimator fitted onto $n$ simulated standard Normal data. Boxplots represent the variation of the error from 100 repeats. The red dashed line shows the bandwidth which minimise the fitted spline.}
    \label{fig:inference_BandwidthSelection2_plot}
\end{figure}

A relationship between the optimal bandwidth and $n$ was attempted to be found. It was assumed such a relationship has the linear from
\begin{equation}
  h_{\text{optimal}} = b n^{-1/5} + a
\end{equation}
where $b$ and $a$ are parameters to be fitted. This was chosen to establish a simple relationship with the rules of thumb. Figure \ref{fig:inference_BandwidthSelection2_ruleOfThumb} show the fitting of an identity link Gamma GLM in order to estimate $b$ and $a$. The estimated parameters are shown in Table \ref{table:inference_BandwidthSelection2_ruleOfThumb}.

\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{../figures/inference/BandwidthSelection2_ruleOfThumb.eps}
  \caption{An identity link gamma GLM was fitted onto the relationship between the optimal bandwidth and $n^{-1/5}$ from the experimental data. The dotted lines shows the 95\% prediction interval.}
  \label{fig:inference_BandwidthSelection2_ruleOfThumb}
\end{figure}

\begin{table}
  \centering
  \input{../figures/inference/BandwidthSelection2_bandwidthEstimate.txt}
  \caption{Estimated and standard error of the gradient and intercept from the linear relationship in Figure \ref{fig:inference_BandwidthSelection2_ruleOfThumb}.}
  \label{table:inference_BandwidthSelection2_ruleOfThumb}
\end{table}

The rules of thumb can be improved if a small bias is added to it. This is because an intercept was found when fitting a linear relationship between the optimal bandwidth and $n^{-1/5}$. Adding this bias should improve the performance of the estimator $\widehat{\sigma}_0$. The gradient is similar to the rules of thumb in the literature \citep{sheather2004density}.

The GLM fit in Figure \ref{fig:inference_BandwidthSelection2_ruleOfThumb} appeared to underfit, which suggest there may be a more complicated relationship between the optimal bandwidth and $n$. The method used in fitting the smoothing spline could contribute to the underfitting. For example if the smoothing spline was not smooth enough so that local minima appear, it would add variance to the location of the optimal bandwidth. A basic linear relationship was chosen to avoid overfitting and keep the rule of thumb simple.

In conclusion, a small bias added to the rule of thumb improved the performance of the estimator $\widehat{\sigma}_0$. \cite{silverman1986density} pointed out that a smaller bandwidth should be used when in particular the distribution is bimodal, this is why \cite{silverman1986density} suggest the use of $b=0.9$ rather than $b=1.06$. Following from this, it was chosen that the intercept was added to Silverman's rule of thumb for the use in the empirical null
\begin{equation}
  h = \left(0.9n^{-1/5}+ 0.16\right) \times\text{min}\left(s_z,\text{IQR}_z/1.34\right) 
  \ .
  \tag{\ref{eq:inference_ourruleofthumb}}
\end{equation}

\afterpage{\clearpage}
\subsection{Sampling Distribution Experiments}

Various estimators of central tendency and dispersion were compared with the empirical null estimates of the null mean and null standard deviation. The empirical sampling distribution of these estimators were observed to look for properties such as bias, variance and robustness. The central tendency looked at were the median and mean. The dispersions looked at were the median absolute deviation (MAD) around the mode $\times 1.4826$, interquartile range (IQR) $\div 1.349$ and the standard deviation. The normalisation constants were used so that the MAD and IQR can be compared directly to the standard deviation.

A dataset of $n = \pi r^2$ standard Normal random variables were simulated to obtain estimates of central tendency and dispersion. This was repeated 100 times to obtain a sampling distribution. The results are shown in Figures \ref{fig:inference_NullIidMean} and \ref{fig:inference_NullIidStd}. The null mean and standard deviation estimations both suffer from high variance, much higher than typical robust estimators such as the median and IQR.

The estimation of the null standard deviation has some positive bias and it was found that this bias is quite sensitive to the bandwidth. Further fine tuning on the bandwidth would be required for the empirical null standard deviation to be unbiased. For example it appeared there is consistent positive bias for all $r$ investigated, perhaps suggesting the bandwidth parameter $a$ should be decreased ever so slightly. The source of bias could come from the fact that the bandwidth was tuned to minimise the mean squared error rather than the squared bias.

To test for robustness, a dataset of $n = \pi r^2$ mixture of Gaussian random variables were simulated with the following distribution: $\normal(0,1)$ with probability $90\%$, $\normal(3,1)$ with probability $10\%$. The $\normal(3,1)$ component acts as non-null statistics and robust estimators should be unaffected by it. The results are shown in Figures \ref{fig:inference_NullIidMixtureMean} and \ref{fig:inference_NullIidMixtureStd}. The empirical null estimators are robust to the non-null component but still suffers from high variance. The rest of the estimators were affected by the non-null statistics, making them unsuitable to estimate parameters of the null component.

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidEmpirical_nullMean.eps}
        \caption{Empirical null mean}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMedianIqr_nullMean.eps}
        \caption{Median}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMeanVar_nullMean.eps}
        \caption{Mean}
      \end{subfigure}
    }
  \caption{Different types of central tendency estimators were used on $\pi r^2$ standard Normal random variables. Shown are the sampling distributions from 100 repeats. The dotted lines show the 95\% confidence interval for the sample mean.}
  \label{fig:inference_NullIidMean}
\end{figure}

\begin{figure}
  \centering
    \centerline{
      \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidEmpirical_nullStd.eps}
        \caption{Empirical null std.}
      \end{subfigure}
      \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMadMode_nullStd.eps}
        \caption{MAD mode}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMedianIqr_nullStd.eps}
        \caption{IQR}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMeanVar_nullStd.eps}
        \caption{Standard deviation}
      \end{subfigure}
    }
  \caption{Different types of statistical dispersion estimators were used on $\pi r^2$ standard Normal random variables. Shown are the sampling distributions from 100 repeats. The dotted lines show the 95\% confidence interval for the sample variance.}
  \label{fig:inference_NullIidStd}
\end{figure}

\begin{figure}
  \centering
    \centerline{
      \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixtureEmpirical_nullMean.eps}
        \caption{Empirical null mean}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixtureMedianIqr_nullMean.eps}
        \caption{Median}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixtureMeanVar_nullMean.eps}
        \caption{Mean}
      \end{subfigure}
    }
  \caption{Different types of central tendency estimators were used on $\pi r^2$ mixture of Gaussian random variables. The random variable has distribution $\normal(0,1)$ with probability $0.9$ and $\normal(3,1)$ with probability $0.1$. Shown are the sampling distributions from 100 repeats. The dotted lines show the 95\% confidence interval for the sample mean.}
  \label{fig:inference_NullIidMixtureMean}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixtureEmpirical_nullStd.eps}
        \caption{Empirical null std.}
      \end{subfigure}
      \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixtureMadMode_nullStd.eps}
        \caption{MAD mode}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixtureMedianIqr_nullStd.eps}
        \caption{IQR}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixtureMeanVar_nullStd.eps}
        \caption{Standard deviation}
      \end{subfigure}
    }
  \caption{Different types of statistical dispersion estimators were used on $\pi r^2$ mixture of Gaussian random variables. The random variable has distribution $\normal(0,1)$ with probability $0.9$ and $\normal(3,1)$ with probability $0.1$. Shown are the sampling distributions from 100 repeats. The dotted lines show the 95\% confidence interval for the sample variance.}
  \label{fig:inference_NullIidMixtureStd}
\end{figure}

\afterpage{\clearpage}
\section{Empirical Null Filter}

Returning back to the example at the start of the chapter, the empirical null could be used on the histogram shown in Figure \ref{fig:inference_inferenceIntro_histogram}. The resulting empirical null distribution is shown in Figure \ref{fig:inference_inferenceSubsampleAll}. The empirical null distribution has a very large standard deviation and would test all pixels as negative, this is incorrect.

\begin{figure}[hpt]
  \centering
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/inferenceSubsampleAll_subimagePositive.eps}
    \caption{$z$ statistics}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/inferenceSubsampleAll_densityEstimate.eps}
    \caption{Empirical null fitted onto histogram}
  \end{subfigure}
  \caption{The resulting $z$ statistics are shown in a). The density estimate of the statistics is shown in b). The empirical null distribution multiplied by some constant is also plotted in b); this is to illustrate the curvature is the same as the density estimate at the mode.}
  \label{fig:inference_inferenceSubsampleAll}
\end{figure}

The problem is that the empirical distribution of the $z$ statistics did not look Normal. It appeared that the $z$ statistics varied spatially, suggesting the empirical null varied spatially as well. One possible reason is because it looked like the value of the $z$ statistics depended on which surface it was on for example.

One possible extension to the empirical null is to spilt the $z$ image into a grid and conduct inference in each section separately. There are problems with this though. The first problem is that how the grid is overlaid can be arbitrary, for example the grid can be translated to produce different sections. Secondly it is not clear how to combine the results from each section together \citep{efron2008simultaneous}.

The empirical null filter extends the empirical null to images. The parameters of the null distribution varies spatially, slowly and smoothly in 2D space. Let the test statistic of the pixel at position $(x,y)$ be
\begin{equation}
Z_{x,y}\sim\normal(\mu_{x,y},\widehat{\sigma}_{x,y}^2)
\end{equation}
for $x=1,2,\cdots,W$ and $y=1,2,\cdots,H$. $\mu_{0,x,y}$ and $\sigma_{0,x,y}$ are the null mean and null standard deviation at position $(x,y)$ respectively. Define the null hypotheses to be
\begin{equation}
  H_{0,x,y}:\mu_{x,y}=\widehat{\mu}_{0,x,y}
\end{equation}
which are tested against
\begin{equation}
  H_{1,x,y}:\mu_{x,y}\neq\widehat{\mu}_{0,x,y} \ .
\end{equation}
 The empirical null filter aims to estimate the parameters $\mu_{0,x,y}$ and $\sigma_{0,x,y}$ for all $x$ and $y$.

The empirical null filter uses the empirical null on neighbouring pixels to estimate the null distribution parameters. To estimate $\mu_{0,x,y}$ and $\sigma_{0,x,y}$, a circular kernel of radius $r$ was centred at $(x,y)$. All the pixels captured by the circular kernel were then used for the empirical null to obtain $\widehat{\mu}_{0,x,y}$ and $\widehat{\sigma}_{0,x,y}$, in other words
\begin{equation}
  \widehat{\mu}_{0,x,y}
  =
  \widehat{\mu}_{0,x,y}
    \left(
      \text{all }Z_{i,j}\text{ such that }(x-i)^2+(y-j)^2\leqslant r^2
    \right)
\end{equation}
and
\begin{equation}
  \widehat{\sigma}_{0,x,y}
  =
  \widehat{\sigma}_{0,x,y}
    \left(
      \text{all }Z_{i,j}\text{ such that }(x-i)^2+(y-j)^2\leqslant r^2
    \right)
  \ .
\end{equation}
Using these estimates, the test statistics were normalised using
\begin{equation}
  T_{x,y} = 
  \dfrac{
    Z_{x,y}-\widehat{\mu}_{0,x,y}
  }
  {
    \widehat{\sigma}_{0,x,y}
  }
\end{equation}
and only the normalised test statistics were used in hypothesis testing.

The parameter of choice is the kernel radius $r$. A $r$ too small will treat defects as the null and will not detect them. A $r$ too big can also cause problems as well. The empirical null filter assumes that the null parameters vary smoothly and slowly so that the null test statistics captured by the kernel is somewhat Normal. If $r$ is too big, the kernel will capture far away test statistics with a very different null distribution. If all of these test statistics are pulled together into a histogram, the null distribution may not look Normal at all which can cause problems when fitting a Normal null distribution. Furthermore, if $r\rightarrow \infty$, then this is just using the empirical null on all pixels.

$r$ does control the proportion of pixels which are null captured by the kernel, denoted as $\pi_0$. If the size of the defects are known, then $r$ should be selected such that
\begin{equation}
\dfrac{
  \text{area of defect}
}
{
  \pi r^2
}
<
5\% \text{ to } 10 \%
\end{equation}
so that $\pi_0$ is of a sensible value. In practice, the defects are unknown so various kernel radiuses should be tried out. But in general, the kernel radius should be much larger than the size of the defect.

There exist similar filters, for example the local normalization filter \citep{sage2003teaching, sage2018local} uses a Gaussian filter to obtain $\widehat{\mu}_{0,x,y}$. $\widehat{\sigma}_{0,x,y}^2$ was obtained by applying another Gaussian filter on the $\left(z_{x,y}-\widehat{\mu}_{0,x,y}\right)^2$ image. The purpose of this filter was to normalise images with uneven illumination. The disadvantage of using the local normalization filter for estimating the null parameters is that it uses a Gaussian filter. This is a type weighted mean which won't be robust to non-null statistics. In addition, the two Gaussian filters can have different radiuses, it is not clear how to set them.

\subsection{Design and Implementation}

The empirical null filter was implemented in an open source software called \texttt{ImageJ} \citep{abramoff2004image, schneider2012nih, perez2013image} by modifying the existing class \texttt{RankFilters}, available on \texttt{GitHub} \citep{imagej2018rank}. This class implemented filters, such as the mean filter and median filter, using a circular kernel, making this a suitable framework for the empirical null filter. Figure \ref{fig:inference_fijiGui} shows the graphical user interface of the empirical null filter in \texttt{Fiji} \citep{schindelin2012fiji}, a distribution of \texttt{ImageJ}.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{../figures/inference/fiji/gui.png}
    \caption{Graphical user interface of the empirical null filter in \texttt{Fiji}. The user can adjust the kernel radius as well as other advanced options relating to the Newton-Raphson method and the kernel density estimate. By-products such as the empirical null parameters can be shown after the filtering as well.}
    \label{fig:inference_fijiGui}
\end{figure}

As a test, the empirical null filter with $r=5$ was used on a \texttt{ImageJ} sample image \texttt{bridge.gif}, as shown in Figure \ref{fig:inference_fijiBridgeFilter}. It is meaningless to use the empirical null filter on an image not for the purpose of the filter, however it does verify the filter works on arbitrary images with some computational considerations. The empirical null images $\widehat{\mu}_{0,x,y}$ and $\widehat{\sigma}_{0,x,y}$ for all ${x,y}$ could be of interest. In particular the empirical null mean image could be interpreted as the result of a mode filter \citep{griffin2000mean}. Figure \ref{fig:inference_fijiMean} compares the empirical null mean with other averaging filters. The empirical null mean creates an impasto look and preserve edges which is similar to the result in \cite{griffin2000mean}. Spread filters such as the standard deviation filter can be used to detect edges. Figure \ref{fig:inference_fijiStd} compares the standard deviation filter with the empirical null standard deviation. The resulting images are similar but it was notable that the edges in the empirical null standard deviation are sharper, suggesting that the empirical null standard deviation is a more robust measure of spread compared to the standard deviation.

\begin{figure}
  \centering
    \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/fiji/bridge.png}
        \caption{Before filtering}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/fiji/bridgeFilter.png}
        \caption{After filtering}
    \end{subfigure}
    }
    \caption{Empirical null filter with $r=5$ on the sample image \texttt{bridge.gif}.}
    \label{fig:inference_fijiBridgeFilter}
\end{figure}

\begin{figure}
  \centering
    \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/fiji/bridge.png}
        \caption{No filter}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/fiji/mean.png}
        \caption{Mean filter}
    \end{subfigure}
    }
    \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/fiji/median.png}
        \caption{Median filter}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/fiji/nullMean.png}
        \caption{Empirical null mean}
    \end{subfigure}
    }
    \caption{Various average filters with kernel radius $r=5$ on the sample image \texttt{bridge.gif}}
    \label{fig:inference_fijiMean}
\end{figure}

\begin{figure}
  \centering
    \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/fiji/std.png}
        \caption{Standard deviation filter}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/fiji/nullStd.png}
        \caption{Empirical null standard deviation}
    \end{subfigure}
    }
    \caption{Various spread filters with kernel radius $r=5$ on the sample image \texttt{bridge.gif}.}
    \label{fig:inference_fijiStd}
\end{figure}

When using the the empirical null filter in \texttt{ImageJ}, the user is presented with a menu, as shown in Figure \ref{fig:inference_fijiGui}. Here the user can input the kernel radius, the filter then modifies the currently selected image $z_{x,y}$ to $t_{x,y}$ for all $x$ and $y$ using a circular kernel with the specified radius. The menu also has options for by-product images to be shown such as: the empirical null mean $\widehat{\mu}_{0,x,y}$, the empirical null standard deviation $\widehat{\sigma}_{0,x,y}$, the standard deviation filter, and quartile filters. Advanced options are available to the user. The number of initial values set the number of valid solutions to be found in the Newton-Raphson method on finding the mode of the density estimate, the solution with the gradient closest to zero is then used in further calculations. The user can also set the maximum number of steps and tolerance for the Newton-Raphson method. The bandwidth parameters for the kernel density estimate can also be adjusted here.

Because the empirical null filter was implemented by modifying the existing code in the class \texttt{RankFilters}, most of the design has already been done. \texttt{RankFilters} contain implementations of filters such as the mean filter and median filter. The implementation makes use of multiple threads, each thread filters a row in parallel. This is done by the following, a thread copies the value of each pixels in its current and neighbouring rows into a cache. Then for each column, the thread does calculations using the cache then replaces the pixel value of the image to be filtered. Once the row has been filtered, the thread then moves onto the next unfiltered row. Rows in the cache not needed are overwritten with more recent rows.

When the kernel captures pixels outside the boundary of the image or region of interest (ROI), \texttt{RankFilters} uses nearest pixel padding which fills in pixels outside the ROI with values to the nearest pixels. Figure \ref{fig:inference_padding_nearest} shows an example of nearest pixel padding on the top left corner of a rectangle ROI. This is not suitable for the empirical null filter as this will cause bias in a lot of calculations, especially the density fitting. The empirical null filter uses \texttt{NaN} padding which fills pixels outside the ROI with \texttt{NaN}. This indicate that these pixels are outside the ROI and are ignored. The number of non-\texttt{NaN} captured by the kernel is kept track for calculations such as the bandwidth for the kernel density estimate.

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.49\textwidth}
      \begin{tabular}{lllllll}
         $\ddots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\udots$ \\
         $\cdots$ & $z_{1,1}$ & $z_{1,1}$ & $z_{1,1}$ & $z_{1,2}$ & $\cdots$ \\
         $\cdots$ & $z_{1,1}$ & $z_{1,1}$ & $z_{1,1}$ & $z_{1,2}$ & $\cdots$ \\ \cline{4-6}
         $\cdots$ & $z_{1,1}$ & \multicolumn{1}{l|}{$z_{1,1}$} & $z_{1,1}$ & $z_{1,2}$ & $\cdots$ \\
         $\cdots$ & $z_{2,1}$ & \multicolumn{1}{l|}{$z_{2,1}$}  & $z_{2,1}$ & $z_{2,2}$ & $\cdots$ \\
         $\udots$ & $\vdots$ & \multicolumn{1}{l|}{$\vdots$}  & $\vdots$ & $\vdots$ & $\ddots$
    \end{tabular}
    \caption{Nearest pixel padding}
    \label{fig:inference_padding_nearest}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
      \begin{tabular}{lllllll}
         $\ddots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\udots$ \\
         $\cdots$ & \texttt{NaN} & \texttt{NaN} & \texttt{NaN} & \texttt{NaN} & $\cdots$ \\
         $\cdots$ & \texttt{NaN} & \texttt{NaN} & \texttt{NaN} & \texttt{NaN} & $\cdots$ \\ \cline{4-6}
         $\cdots$ & \texttt{NaN} & \multicolumn{1}{l|}{\texttt{NaN}} & $z_{1,1}$ & $z_{1,2}$ & $\cdots$ \\
         $\cdots$ & \texttt{NaN} & \multicolumn{1}{l|}{\texttt{NaN}}  & $z_{2,1}$ & $z_{2,2}$ & $\cdots$ \\
         $\udots$ & $\vdots$ & \multicolumn{1}{l|}{$\vdots$}  & $\vdots$ & $\vdots$ & $\ddots$
    \end{tabular}
    \caption{\texttt{NaN} padding}
    \label{fig:inference_padding_nan}
  \end{subfigure}
  \caption{When a kernel contain pixels outside the region of interest as shown by the solid line, the missing pixels can either be extrapolated using the nearest pixel or completely ignored by filling in the missing pixels with \texttt{NaN}.}
  \label{fig:inference_padding}
\end{figure}

Line filtering is done from left to right. At the start on the far left, an initial value of the median of all the pixels in the kernel is used. In addition, various initial values were randomly tried out until a requested number of valid solutions are found, the solution with the greatest value of $p_Z(\widehat{\mu}_0)$ was used. For the following pixel to the right, the empirical null mean of the neighbouring left pixel is used as a initial point. This was chosen as it was assumed the empirical null mean would vary slowly and smoothly spatially.

Trying out more initial values is an advantage because it eliminates the chances of obtaining a sub-optimal solution such as a local maxima. However this is traded for computational cost. Additional initial values were generated by sampling from the Normal distribution centred on the first initial value and with standard deviation the same as the standard deviation using all pixels in the kernel.

When filtering arbitrary images such as \texttt{bridge.gif}, a few computational considerations were needed. \texttt{bridge.gif} is an 8-bit image so values are represented as integers. This open up opportunities for the inter-quartile range and the standard deviation to be zero, this caused problems as this set the bandwidth to be zero. When the standard deviation is zero, it was set to $0.289$ which corresponds to the standard deviation of a Uniform random variable. When the inter-quartile range is zero, the inter-quartile range was set to the standard deviation $\times 1.34$.

When the number of initial values is too few, poor solutions to $\widehat{\mu}_0$ are uncommon which then propagate to the pixel to the right and used as its initial value. Poor initial values could cause the Newton-Raphson method to never converge. If the Newton-Raphson method failed to converge too many times, the median is used as the new initial value. Should it fail too, the program gives up and puts a $\texttt{NaN}$ in that pixel.

\subsection{All-null Experiment}

An experiment was conducted to demonstrate the effects of the empirical null filter on a simulated $256\times 256$ Gaussian image, where all the pixels have value distributed as standard Normal. The purpose is to simulate test statistics which are null and ideally no positive results should be detected. The test statistics after filtering should preserve its original distribution, or at least close to it. The statistical moments before and after filtering should be the same and were investigated. Because the empirical null mean and empirical null standard deviation are random variables with some distribution, the filtered pixels will never be Normal and any Normality tests would be too strict for the purpose of this experiment.

\emph{Contamination} is defined as some smooth and slowly varying function added and/or multiplied to an image. When contaminating a Gaussian image, it may cause errors in hypothesis testing if the null distribution was incorrectly specified as standard Normal. In this experiment, the contamination was a gradient such that the pixel at position $(x,y)$ have a value distributed as $Z_{x,y}\sim\normal(\mu_{0,x,y},\sigma_{0,x,y}^2)$ where
\begin{equation}
  \mu_{0,x,y} = 0.01 (x-x_0) + 0.01 (y-y_0) \ ,
\end{equation}
$(x_0,y_0)$ is the centre of the image and $\sigma_{0,x,y}=2$ for all $(x,y)$.

A caveat is that some bias to the null variance estimation would be introduced because of the contamination. This is because sources of variance captured by the kernel are from $\sigma_{0,x,y}$ and also from the variability of $\mu_{0,x,y}$. This can be shown. Let $Z_{x,y}$ be the value of the pixels captured by the circular kernel for integer values of $(x,y)$ such that $x^2+y^2\leqslant r$. Suppose that $Z_{x,y}\sim\normal(ax+by,c^2\sigma^2)$. The calculations of the expectation and variance of $Z_{x,y}$ can be approximated by treating $x$ and $y$ as uniformly distributed within a circle centred at the origin with radius $r$. The expectation is $\expectation[Z_{X,Y}] = \expectation\expectation[Z_{X,Y}|X,Y]=0$. The variance is given as $\variance[Z_{X,Y}] = \expectation\variance[Z_{X,Y}|X,Y] + \variance\expectation[Z_{X,Y}|X,Y]$. From the distribution of $Z_{X,Y}$, $\expectation\variance[Z_{X,Y}|X,Y] = c^2\sigma^2$ and $\variance\expectation[Z_{X,Y}|X,Y]=\variance[aX+bY]$. This is worked out to be
\begin{align}
\variance[aX+bY] &= \dfrac{1}{\pi r^2}\int_{\rho=0}^{\rho=r}\int_{\theta=0}^{\theta=2\pi}
\rho^3(a\cos\theta + b\sin\theta)^2 \diff \theta \diff \rho
\nonumber\\
& = \dfrac{1}{4}(a^2+b^2)r^2
\end{align}
thus
\begin{equation}
\variance[Z_{X,Y}] = \dfrac{1}{4}(a^2+b^2)r^2 + c^2\sigma^2 \ .
\end{equation}
This shows that when estimating the null variance on a contaminated Gaussian image, there may be a $(a^2+b^2)r^2/4$ bias. If the contamination is slowly varying, that is $a$ and $b$ are small, then the bias should be small.

\begin{figure}[p]
  \centering
  \centerline{
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/allNullGaussianScript_beforeFilter.eps}
    \caption{Unfiltered}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/allNullGaussianScript_afterFilter.eps}
    \caption{Filtered}
    \label{fig:inference_allNullGaussianScript_afterFilter}
  \end{subfigure}
  }
  \centerline{
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/allNullGaussianScript_nullMean.eps}
    \caption{Empirical null mean}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/allNullGaussianScript_nullStd.eps}
    \caption{Empirical null std}
  \end{subfigure}
  }
  \caption{A $256\times 256$ Gaussian image before and after filtering with kernel radius 20. Also shown are the empirical null mean and empirical null std.}
  \label{fig:inference_allNullGaussianScript}
\end{figure}

\begin{figure}[p]
  \centering
  \centerline{
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/allNullPlaneScript_beforeFilter.eps}
    \caption{Unfiltered}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/allNullPlaneScript_afterFilter.eps}
    \caption{Filtered}
    \label{fig:inference_allNullPlaneScript_afterFilter}
  \end{subfigure}
  }
  \centerline{
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/allNullPlaneScript_nullMean.eps}
    \caption{Empirical null mean}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/allNullPlaneScript_nullStd.eps}
    \caption{Empirical null std}
  \end{subfigure}
  }
  \caption{A $256\times 256$ contaminated Gaussian image before and after filtering with kernel radius 20. Also shown are the empirical null mean and empirical null std.}
  \label{fig:inference_allNullPlaneScript}
\end{figure}

\begin{figure}[htp]
  \centering
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/allNullGaussianScript_pValue.eps}
    \caption{No contamination}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/allNullPlaneScript_pValue.eps}
    \caption{With contamination}
  \end{subfigure}
  \caption{$p$-values of values from a filtered $256\times 256$ Gaussian image with kernel radius 20. In b), the Gaussian image was contaminated before filtering. The $p$-values were obtained from the images in Figures \ref{fig:inference_allNullGaussianScript_afterFilter} and \ref{fig:inference_allNullPlaneScript_afterFilter}. The dotted lines shows the $p$-values if they were uniformly distributed.}
  \label{fig:inference_allNull_pValues}
\end{figure}

An example of a Gaussian image with/without contamination before/after filtering are shown in Figures \ref{fig:inference_allNullGaussianScript} and \ref{fig:inference_allNullPlaneScript}. In the contaminated example. the filter managed to estimate the null mean, picking up the gradient. This enabled the normalised image to look flat to remove the contamination. The $p$-values after filtering are shown in Figure \ref{fig:inference_allNull_pValues}. A quick inspection suggest that the filtered pixels looked reasonably Normal for this particular example.

For a given kernel radius $r$, 100 different Gaussian images were filtered to investigate the within image mean, standard deviation and kurtosis of the filtered statistics. Various other filters were used as well such as the MAD mode null filter, median IQR null filter and the mean variance null filter. They use different estimators to estimate the null parameters as their name suggest, Table \ref{table:inference_nullFilters} clarifies them.

\begin{table}[ht]
  \centering
  \begin{tabular}{c|c|c}
    Filters&Null mean&Null Standard deviation\\\hline
    MAD mode null&Mode&MAD around the mode $\times1.4826$\\
    Median IQR null&Median&IQR $\div1.349$\\
    Mean variance null&Mean&Standard deviation
  \end{tabular}
  \caption{Various filters using different estimators to estimate the null parameters are described here.}
  \label{table:inference_nullFilters}
\end{table}

\begin{figure}[htp]
  \centering
  \centerline{
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianEmpirical_mean.eps}
    \caption{Empirical Null}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianMadMode_mean.eps}
    \caption{MAD Mode Null}
  \end{subfigure}
  }
  \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianMedianIqr_mean.eps}
    \caption{Median IQR Null}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianMeanVar_mean.eps}
    \caption{Mean Var Null}
  \end{subfigure}
  }
  \caption{The within image sample \emph{mean} of the test statistics in a filtered Gaussian image of size $256\times256$. The box plots summarise the 100 repeats of the experiment. The dashed lines show the 95\% confidence interval using standard tests, assuming independence.}
  \label{fig:inference_AllNullGaussian_mean}
\end{figure}

\begin{figure}[htp]
  \centering
  \centerline{
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianEmpirical_variance.eps}
    \caption{Empirical Null}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianMadMode_variance.eps}
    \caption{MAD Mode Null}
  \end{subfigure}
  }
  \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianMedianIqr_variance.eps}
    \caption{Median IQR Null}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianMeanVar_variance.eps}
    \caption{Mean Var Null}
  \end{subfigure}
  }
  \caption{The within image sample \emph{standard deviation} of the test statistics in a filtered Gaussian image of size $256\times256$. The box plots summarise the 100 repeats of the experiment. The dashed lines show the 95\% confidence interval using standard tests, assuming independence.}
  \label{fig:inference_AllNullGaussian_var}
\end{figure}

\begin{figure}[htp]
  \centering
  \centerline{
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianEmpirical_kurtosis.eps}
    \caption{Empirical Null}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianMadMode_kurtosis.eps}
    \caption{MAD Mode Null}
  \end{subfigure}
  }
  \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianMedianIqr_kurtosis.eps}
    \caption{Median IQR Null}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullGaussianMeanVar_kurtosis.eps}
    \caption{Mean Var Null}
  \end{subfigure}
  }
  \caption{The within image sample \emph{kurtosis} of the test statistics in a filtered Gaussian image of size $256\times256$. The box plots summarise the 100 repeats of the experiment. The dashed lines show the 95\% confidence interval using standard tests, assuming independence.}
  \label{fig:inference_AllNullGaussian_kurtosis}
\end{figure}

\begin{figure}[htp]
  \centering
  \centerline{
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullPlaneEmpirical_mean.eps}
    \caption{Empirical Null}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullPlaneMadMode_mean.eps}
    \caption{MAD Mode Null}
  \end{subfigure}
  }
  \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullPlaneMedianIqr_mean.eps}
    \caption{Median IQR Null}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullPlaneMeanVar_mean.eps}
    \caption{Mean Var Null}
  \end{subfigure}
  }
  \caption{The within image sample \emph{mean} of the test statistics in a filtered \emph{contaminated} Gaussian image of size $256\times256$. The box plots summarise the 100 repeats of the experiment. The dashed lines show the 95\% confidence interval using standard tests, assuming independence.}
  \label{fig:inference_AllNullPlane_mean}
\end{figure}

\begin{figure}[htp]
  \centering
  \centerline{
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullPlaneEmpirical_variance.eps}
    \caption{Empirical Null}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullPlaneMadMode_variance.eps}
    \caption{MAD Mode Null}
  \end{subfigure}
  }
  \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullPlaneMedianIqr_variance.eps}
    \caption{Median IQR Null}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullPlaneMeanVar_variance.eps}
    \caption{Mean Var Null}
  \end{subfigure}
  }
  \caption{The within image sample \emph{standard deviation} of the test statistics in a filtered \emph{contaminated} Gaussian image of size $256\times256$. The box plots summarise the 100 repeats of the experiment. The dashed lines show the 95\% confidence interval using standard tests, assuming independence.}
  \label{fig:inference_AllNullPlane_var}
\end{figure}

\begin{figure}[htp]
  \centering
  \centerline{
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullPlaneEmpirical_kurtosis.eps}
    \caption{Empirical Null}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullPlaneMadMode_kurtosis.eps}
    \caption{MAD Mode Null}
  \end{subfigure}
  }
  \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullPlaneMedianIqr_kurtosis.eps}
    \caption{Median IQR Null}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{../figures/inference/AllNullPlaneMeanVar_kurtosis.eps}
    \caption{Mean Var Null}
  \end{subfigure}
  }
  \caption{The within image sample \emph{kurtosis} of the test statistics in a filtered \emph{contaminated} Gaussian image of size $256\times256$. The box plots summarise the 100 repeats of the experiment. The dashed lines show the 95\% confidence interval using standard tests, assuming independence.}
  \label{fig:inference_AllNullPlane_kurtosis}
\end{figure}

The results for the filtered images are shown in Figures \ref{fig:inference_AllNullGaussian_mean}, \ref{fig:inference_AllNullGaussian_var} and \ref{fig:inference_AllNullGaussian_kurtosis}. For the filtered \emph{contaminated} images, they are in Figures \ref{fig:inference_AllNullPlane_mean}, \ref{fig:inference_AllNullPlane_var} and \ref{fig:inference_AllNullPlane_kurtosis}. The filtered test statistic means did agree with zero as expected, showing that the filters centred the statistics when normalising. When using the empirical null filter, there was some negative bias with the standard deviation of the filtered test statistics. The negative bias comes from the positive bias when estimating the null standard deviation.

When filtering the contaminated image, all estimators suffered from negative bias in the standard deviation of the filtered statistics, in particular when the kernel radius increases. This is shown in Figure \ref{fig:inference_AllNullPlane_var}. As discussed before, the kernel captures the variation due to the contamination and this adds intrinsic bias to the null variance. As a result, this adds bias when estimating the null variance.

The kurtosis in Figures \ref{fig:inference_AllNullGaussian_kurtosis} and \ref{fig:inference_AllNullPlane_kurtosis} shows that the empirical null filter can cause the filtered statistics to have heavy tails, in particular for small kernel radiuses. This could cause problems in hypothesis testing because heavy tails could be misinterpreted as contribution from non-null statistics. The source of the kurtosis inflation is from the estimation of the null standard deviation. This can be seen by comparing the empirical null filter with the MAD mode null filter because the only difference between the two filters is in the estimation of the null standard deviation.

\afterpage{\clearpage}
\subsection{Detecting Defects Experiment}

Figure \ref{fig:inference_contamination_images} shows an example of a contaminated image before and after filtering. When doing a hypothesis test, the contamination caused systematic error such that there are false positives on the two corners of the unfiltered image. The empirical null filter can recover the original image by extracting the systematic errors such as the plane, as shown by the empirical null mean, and the increase of variance, as shown by the empirical null std.

The QQ plot of the filtered test statistics, as shown in Figure \ref{fig:inference_contamination_qq}, shows that there may be a variance mismatch. A mismatch in the variance will increase or decrease the sensitivity of the test, a larger variance would increase the sensitivity.

An experiment was conducted to investigate the sample mean and sample variance of the filtered contaminated image. Ideally the sample mean and sample variance should be 0 and 1 respectively. Different radiuses were investigated from ??? to ??? and repeated ??? times.

The results are shown in Figure \ref{fig:inference_allnullplane}. The filter correctly centred the test statistics at zero. However there was some variance mismatch and this depended on the kernel radius. A kernel radius of about 80 scaled the test statistics correctly. This suggest there may be an optimal kernel radius for this particular null distribution. Another caveat is that the sensitivity may not be exactly preserved after filtering.

The study will now turn to contaminated defected images. Let $Z_{x,y}$ be the test statistics of pixel $(x,y)$ and under the null, it is distributed such that $Z_{x,y}|H_0\sim\normal(0,1)$. Defects refer to statistics which do not follow the null distribution, but instead an alternative distribution. For example it can be $Z_{x,y}|H_1\sim\normal(\mu_{1,x,y},1)$ and this example was used in this experiment. Contamination is a linear transform of the test statistics $Z_{x,y}$ such that
\begin{equation}
  C_{x,y} = Z_{x,y}\sigma_{0,x,y} + \mu_{0,x,y} \ .
\end{equation}
The empirical null filter recovers $Z_{x,y}$ from $C_{x,y}$. The values of the filtered test statistics are denoted as $\zeta_{x,y}$.

Various simulated defects were investigated. Salt defect (named after salt and pepper noise) with density $\pi_1$ assign all test statistics to be distributed such that
\begin{equation}
  p_{Z_{x,y}}(z)=(1-\pi_1)p_{Z_{x,y}|H_0}(z) + \pi_1 p_{Z_{x,y}|H_1}(z) \ .
\end{equation}
In other words, test statistics will have the alternative distribution with probability $\pi_1$, and have the null distribution otherwise. This was chosen because the proportion of null statistics captured by a kernel should be independent of the radius of the kernel.

A line defect assign columns of pixels to have the alternative distribution, a kernel would only capture a section of defect. Similarly different shapes were used such as a square defect. With a square defect, a kernel can capture the entire defect.

An example of a salt defected image is shown in Figure \ref{fig:inference_saltDefect}, the figure also includes the image contaminated and filtered after contamination. The contamination caused some systematic errors, such as increasing the sensitivity such that are there more positive pixels in particular in two of the corners of the image. The empirical null filter recovers the defected image from the contaminated image so that appropriate inference can be done. However some statistical power is lost in the filtered image. This is because in areas of high empirical null std, the sensitivity decreased thus there is less detection power. This cannot be avoided as areas of high empirical null std is due to the sampling error. Using different kernel radiuses will change the sampling error of the empirical null std but it cannot be eliminated. Similar comments can be made for the line defect as shown in Figure \ref{fig:inference_lineDefect}.

The square defected images, including the contaminated ones, are shown in Figures \ref{fig:inference_squareDefect} and \ref{fig:inference_square2Defect}. Figure \ref{fig:inference_squareDefect} filtered the contaminated image using a kernel radius of $r=20$, while Figure \ref{fig:inference_square2Defect} used a kernel radius of $r=40$. With $r=20$, the kernel is smaller than the square defect, therefore the defect was treated as the null and this shows in the empirical null mean. This resulted in difficulty detecting the defect. With $r=40$, the kernel is much bigger than the defect and treated the defect as non-null. The empirical null filter recovers the gradient in the empirical null which was then used for inference.

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/defectDustExample_imagePreContaminated.eps}
        \caption{Uncontaminated}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/defectDustExample_imageContaminated.eps}
        \caption{Contaminated}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/defectDustExample_imageFiltered.eps}
        \caption{Filtered}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/defectDustExample_nullStd.eps}
        \caption{Empirical null std}
      \end{subfigure}
    }
  \caption{A $256 \times 256$ dust defect Gaussian image with/without contamination and after filtering with kernel radius 20. Defected pixels have the distribution $\normal(3,1)$. The salt defect has density $\pi_1 = 0.1$. The contamination is such that the null distribution is $Z_{x,y}\sim\normal(\mu_{0,x,y},\sigma_0^2)$ where $\mu_{0,x,y} = 0.01 (x-x_0) + 0.01 (y-y_0)$ and $\sigma_0=2$. Highlighted in red are significant pixels when controlling the FDR at the $2\sigma$ level.}
  \label{fig:inference_saltDefect}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/defectLineExample_imagePreContaminated.eps}
        \caption{Uncontaminated}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/defectLineExample_imageContaminated.eps}
        \caption{Contaminated}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/defectLineExample_imageFiltered.eps}
        \caption{Filtered}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/defectLineExample_nullStd.eps}
        \caption{Empirical null std}
      \end{subfigure}
    }
  \caption{A $256 \times 256$ line defect Gaussian image with/without contamination and after filtering with kernel radius 20. Defected pixels have the distribution $\normal(3,1)$. The line is 5 pixels thick. The contamination is such that the null distribution is $Z_{x,y}\sim\normal(\mu_{0,x,y},\sigma_0^2)$ where $\mu_{0,x,y} = 0.01 (x-x_0) + 0.01 (y-y_0)$ and $\sigma_0=2$. Highlighted in red are significant pixels when controlling the FDR at the $2\sigma$ level.}
  \label{fig:inference_lineDefect}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/defectSquareExample_imagePreContaminated.eps}
        \caption{Uncontaminated}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/defectSquareExample_imageContaminated.eps}
        \caption{Contaminated}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/defectSquareExample_imageFiltered.eps}
        \caption{Filtered}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/defectSquareExample_nullMean.eps}
        \caption{Empirical null mean}
      \end{subfigure}
    }
  \caption{A $256 \times 256$ square defect Gaussian image with/without contamination and after filtering with kernel radius 20. Defected pixels have the distribution $\normal(3,1)$. The square is $30\times 30$ pixels in size. The contamination is such that the null distribution is $Z_{x,y}\sim\normal(\mu_{0,x,y},\sigma_0^2)$ where $\mu_{0,x,y} = 0.01 (x-x_0) + 0.01 (y-y_0)$ and $\sigma_0=2$. Highlighted in red are significant pixels when controlling the FDR at the $2\sigma$ level.}
  \label{fig:inference_squareDefect}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/defectSquare2Example_imagePreContaminated.eps}
        \caption{Uncontaminated}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/defectSquare2Example_imageContaminated.eps}
        \caption{Contaminated}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/defectSquare2Example_imageFiltered.eps}
        \caption{Filtered}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/defectSquare2Example_nullMean.eps}
        \caption{Empirical null mean}
      \end{subfigure}
    }
  \caption{A $256 \times 256$ square defect Gaussian image with/without contamination and after filtering with kernel radius 40. Defected pixels have the distribution $\normal(3,1)$. The square is $30\times 30$ pixels in size. The contamination is such that the null distribution is $Z_{x,y}\sim\normal(\mu_{0,x,y},\sigma_0^2)$ where $\mu_{0,x,y} = 0.01 (x-x_0) + 0.01 (y-y_0)$ and $\sigma_0=2$. Highlighted in red are significant pixels when controlling the FDR at the $2\sigma$ level.}
  \label{fig:inference_square2Defect}
\end{figure}

The receiver operating characteristic (ROC) curves \citep{green1966signal, metz1978basic, hanley1982meaning, friedman2001elements, cook2007use} are shown in Figure \ref{fig:inference_defectSimulationRoc} for the various defects. The ROC curve is a parametric plot, plotting the true positive rate against the false positive rate for varying thresholds. Other authors \citep{metz1978basic} use terminology such as sensitivity and specificity. The area under the ROC curve is a commonly used statistic to quantify the performance of the test \citep{friedman2001elements}. Interpretations of the area do exist \citep{metz1978basic,hanley1982meaning} and discussed thoroughly in \cite{cook2007use}. The area under the ROC curve was recorded for different intensity values of the defect.

\begin{figure}[htp]
  \centering
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/defectDustExample_roc.eps}
        \caption{Dust defect $\pi_1=0.1$, $r=20$}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/defectLineExample_roc.eps}
        \caption{5 pixels thick line defect, $r=20$}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/defectSquareExample_roc.eps}
        \caption{$30\times 30$ square defect, $r=20$}
        \label{fig:inference_defectSimulationRoc_square20}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/defectSquare2Example_roc.eps}
        \caption{$30\times 30$ square defect, $r=40$}
      \end{subfigure}
    }
  \caption{ROC curves for various defected $256\times 256$ Gaussian images. The difference curves show the results for the non-contaminated image, contaminated image, and filtered contaminated image. Defected pixels have the distribution $\normal(3,1)$. The contamination is such that the null distribution is $Z_{x,y}\sim\normal(\mu_{0,x,y},\sigma_0^2)$ where $\mu_{0,x,y} = 0.01 (x-x_0) + 0.01 (y-y_0)$ and $\sigma_0=2$.}
  \label{fig:inference_defectSimulationRoc}
\end{figure}

In terms of the area under the ROC curves, the empirical null filter improves the performance over using the unfiltered contaminated image. The performance before contamination cannot be recovered but it is an improvement. For kernel radiuses too small, such as Figure \ref{fig:inference_defectSimulationRoc_square20}, the empirical null filter deteriorates the performance dramatically and it would be better off using the contaminated image.

The ROC is useful because it considers all thresholds or sensitivities used in the hypothesis testing. However in the previous experiment, it was found there was some variance mismatch before and after filtering with regards to the test statistics. As a result, for a given threshold, such as controlling the FDR, the sensitivity may change ever so slightly after filtering.

An experiment was conducted to investigate the sensitivity before and after filtering. A $256\times 256$ Gaussian image with dust defect of density $\pi_1=0.1$ was investigated. A kernel radius of $r=20$ was used. The type 1 error, type 2 error and the FDR were investigated before contamination, after contamination and after filtering for various alternative distribution. The alternative distribution used is $Z_{x,y}|H_1\sim\normal(\mu_1,1)$ where $\mu_1$ is the alternative mean and was varied for this experiment.

\begin{figure}[htp]
  \centering
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectAltDust0Empirical_roc.eps}
        \caption{ROC area}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectAltDust0Empirical_type1.eps}
        \caption{Type 1 error}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectAltDust0Empirical_type2.eps}
        \caption{Type 2 error}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectAltDust0Empirical_fdr.eps}
        \caption{FDR}
      \end{subfigure}
    }
  \caption{ROC area and various errors obtained when conducting hypothesis test on a \emph{empirical null} filtered image of a $256 \times 256$ salt defected Gaussian image. a) shows the ROC area when testing the image before and after filtering. In b) c) d) the dotted line shows the 95\% empirical confidence interval of the resulting error when testing the image before filtering. The filter used a kernel radius Defected pixels have the distribution $\normal(\mu_1,1)$ where $\mu_1$ was varied in this experiment. The salt defect has density $\pi_1 = 0.1$. The FDR was controlled at the 5\% level. The box plot summarise the 100 repeats of the experiment.}
\end{figure}

\begin{figure}[htp]
  \centering
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectAltDust0MadMode_roc.eps}
        \caption{ROC area}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectAltDust0MadMode_type1.eps}
        \caption{Type 1 error}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectAltDust0MadMode_type2.eps}
        \caption{Type 2 error}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectAltDust0MadMode_fdr.eps}
        \caption{FDR}
      \end{subfigure}
    }
  \caption{ROC area and various errors obtained when conducting hypothesis test on a \emph{MAD mode null} filtered image of a $256 \times 256$ salt defected Gaussian image. a) shows the ROC area when testing the image before and after filtering. In b) c) d) the dotted line shows the 95\% empirical confidence interval of the resulting error when testing the image before filtering. The filter used a kernel radius  Defected pixels have the distribution $\normal(\mu_1,1)$ where $\mu_1$ was varied in this experiment. The salt defect has density $\pi_1 = 0.1$. The FDR was controlled at the 5\% level. The box plot summarise the 100 repeats of the experiment.}
\end{figure}

\begin{figure}[htp]
  \centering
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectAltDustEmpirical_roc.eps}
        \caption{ROC area}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectAltDustEmpirical_type1.eps}
        \caption{Type 1 error}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectAltDustEmpirical_type2.eps}
        \caption{Type 2 error}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectAltDustEmpirical_fdr.eps}
        \caption{FDR}
        \label{fig:inference_DefectSimulationDust_fdr}
      \end{subfigure}
    }
  \caption{ROC area and various errors obtained when conducting hypothesis test on a \emph{empirical null} filtered contaminated image of a $256 \times 256$ salt defected Gaussian image. a) shows the ROC area when testing the image before/after contamination and after filtering. In b) c) the dotted line shows the 95\% empirical confidence interval of the resulting error when testing the image before contamination. The filter used a kernel radius . Defected pixels have the distribution $\normal(\mu_1,1)$ where $\mu_1$ was varied in this experiment. The salt defect has density $\pi_1 = 0.1$. The contamination is such that the null distribution is $Z_{x,y}\sim\normal(\mu_{0,x,y},\sigma_0^2)$ where $\mu_{0,x,y} = 0.01 (x-x_0) + 0.01 (y-y_0)$ and $\sigma_0=2$. The FDR was controlled at the 5\% level. The box plot summarise the 100 repeats of the experiment.}
  \label{fig:inference_DefectSimulationDust}
\end{figure}

\begin{figure}[htp]
  \centering
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectAltDustMadMode_roc.eps}
        \caption{ROC area}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectAltDustMadMode_type1.eps}
        \caption{Type 1 error}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectAltDustMadMode_type2.eps}
        \caption{Type 2 error}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectAltDustMadMode_fdr.eps}
        \caption{FDR}
        \label{fig:inference_DefectSimulationDust_fdr}
      \end{subfigure}
    }
  \caption{ROC area and various errors obtained when conducting hypothesis test on a \emph{MAD mode null} filtered contaminated image of a $256 \times 256$ salt defected Gaussian image. a) shows the ROC area when testing the image before/after contamination and after filtering. In b) c) the dotted line shows the 95\% empirical confidence interval of the resulting error when testing the image before contamination. The filter used a kernel radius . Defected pixels have the distribution $\normal(\mu_1,1)$ where $\mu_1$ was varied in this experiment. The salt defect has density $\pi_1 = 0.1$. The contamination is such that the null distribution is $Z_{x,y}\sim\normal(\mu_{0,x,y},\sigma_0^2)$ where $\mu_{0,x,y} = 0.01 (x-x_0) + 0.01 (y-y_0)$ and $\sigma_0=2$. The FDR was controlled at the 5\% level. The box plot summarise the 100 repeats of the experiment.}
\end{figure}

The results are shown in Figure \ref{fig:inference_DefectSimulationDust}. The ROC areas were quantified here and shows that, in terms of the ROC area, using the filter improved the performance of the test from the contamination. Figure \ref{fig:inference_DefectSimulationDust_fdr} illustrate that in the uncontaminated image, the FDR is controlled below 5\%. However after contamination and then filtering, the FDR was not well controlled at this threshold. Ideally the error rates before contamination and after filtering should be similar so that sensitivity is preserved after filtering. This should highlight that while performance can be recovered from contamination in terms of the ROC area, the FDR threshold may not be preserved.

There is also the question of what kernel size to choose for a given defect. Literature, such as \cite{efron2004large} and \cite{schwartzman2008empirical}, suggest that, as a rule of thumb, the proportion of non-null statistics should not be larger than $\pi_1=0.1$ to satisfy some assumptions made for the empirical null. Given the size of the defect, one could work out the minimum kernel radius by setting a threshold for the maximum proportion of the area of kernel which is a defect to 10\%. In other words for a kernel with radius $r$
\begin{equation}
0.1 > \dfrac{\text{maximum area of defect captured by the kernel}}{\text{area of kernel}} \ .
\end{equation}
For a $d \times d$ square defect this is
\begin{equation}
r > d \sqrt{\dfrac{10}{\pi}} \ .
\end{equation}
In the example of $d=30$, this is about $r>54$. For the line defect with thickness $d$, it can be involved because the maximum area of the line defect captured by the kernel is
\begin{equation}
2r^2
\left[
  \dfrac{d}{2r}\sqrt{1-\left(\dfrac{d}{2r}\right)^2}
  +\text{arcsin}\left(\dfrac{d}{2r}\right)
\right] \ .
\end{equation}
For $d=5$, the radius is about $r>32$. These are rules of thumb because the kernel won't be perfectly circular when used on a grid of pixels. It should be pointed out that the implementation can accept non-integer radiuses if desired.

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectRadiusLine_roc.eps}
        \caption{ROC area}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectRadiusLine_type1.eps}
        \caption{Type 1 error}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectRadiusLine_type2.eps}
        \caption{Type 2 error}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectRadiusLine_fdr.eps}
        \caption{FDR}
      \end{subfigure}
    }
  \caption{ROC area and various errors obtained when conducting hypothesis test on the filtered contaminated image of a $256 \times 256$ line defected Gaussian image using various kernel radiuses. The dotted lines show the 95\% percentile when the hypothesis test was done on the uncontaminated image. Defected pixels have the distribution $\normal(3,1)$. The line defect is 5 pixels thick. The contamination is such that the null distribution is $Z_{x,y}\sim\normal(\mu_{0,x,y},\sigma_0^2)$ where $\mu_{0,x,y} = 0.01 (x-x_0) + 0.01 (y-y_0)$ and $\sigma_0=2$. The FDR was controlled at the $2\sigma$ level. The box plot summarise the 100 repeats of the experiment.}
  \label{fig:inference_Experiment_DefectRadiusLine}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectRadiusSquare_roc.eps}
        \caption{ROC area}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectRadiusSquare_type1.eps}
        \caption{Type 1 error}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectRadiusSquare_type2.eps}
        \caption{Type 2 error}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectRadiusSquare_fdr.eps}
        \caption{FDR}
      \end{subfigure}
    }
  \caption{ROC area and various errors obtained when conducting hypothesis test on the filtered contaminated image of a $256 \times 256$ square defected Gaussian image using various kernel radiuses. The dotted lines show the 95\% percentile when the hypothesis test was done on the uncontaminated image. Defected pixels have the distribution $\normal(3,1)$. The square defect is $30\times 30$ pixels in size. The contamination is such that the null distribution is $Z_{x,y}\sim\normal(\mu_{0,x,y},\sigma_0^2)$ where $\mu_{0,x,y} = 0.01 (x-x_0) + 0.01 (y-y_0)$ and $\sigma_0=2$. The FDR was controlled at the $2\sigma$ level. The box plot summarise the 100 repeats of the experiment.}
  \label{fig:inference_Experiment_DefectRadiusSquare}
\end{figure}

An experiment was conducted where the contaminated defected image were filtered for various kernel radiuses for a fixed alternative distribution $\normal(3,1)$. Hypothesis testing was done on the uncontaminated and filtered contaminated image. The area under the ROC curve and various errors were recorded. Results for the line and square defect are shown in Figures \ref{fig:inference_Experiment_DefectRadiusLine} and \ref{fig:inference_Experiment_DefectRadiusSquare} respectively.

The results show that the ROC area increased with kernel radius. This does highlight that a kernel with a good radius can performance almost as though there was no contamination. For large kernel radius, the FDR becomes similar to the FDR obtained from the uncontaminated image. Thus it appears a large kernel radius helped preserve FDR control after filtering in these examples. However a large kernel radius does come at a computational cost. Choosing a kernel radius is a matter of trading computational cost with precise FDR control.

\section{Results}

\begin{figure}[htp]
  \centering
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSep16120deg_radius1_z.eps}
        \caption{$r=\addNumber{../figures/inference/DefectDetectSep16120deg_radius1.txt}$}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSep16120deg_radius2_z.eps}
        \caption{$r=\addNumber{../figures/inference/DefectDetectSep16120deg_radius2.txt}$}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSep16120deg_radius3_z.eps}
        \caption{$r=\addNumber{../figures/inference/DefectDetectSep16120deg_radius3.txt}$}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSep16120deg_radius4_z.eps}
        \caption{$r=\addNumber{../figures/inference/DefectDetectSep16120deg_radius4.txt}$}
      \end{subfigure}
    }
  \caption{Filtered $z$ image from the \texttt{Sep16 120deg} dataset. Various kernel radiuses were used.}
  \label{fig:inference_sep16120deg_zfiltered}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSep16120deg_radius1_sig.eps}
        \caption{Image with sig.~pixels}
      \end{subfigure}
      \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSep16120deg_radius1_logp.eps}
        \caption{$-\log p$}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSep16120deg_radius1_nullMean.eps}
        \caption{Empirical null mean}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSep16120deg_radius1_nullStd.eps}
        \caption{Empirical null std}
      \end{subfigure}
    }
  \caption{Resulting inference when using the empirical null filter with radius $r=\addNumber{../figures/inference/DefectDetectSep16120deg_radius1.txt}$. a) Highlighted in red are significant pixels when controlling the FDR at the $2\sigma$ level. b) $p$-values on the log scale. c) and d) Empirical null mean and std.}
  \label{fig:inference_sep16120deg_inference_radius1}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSep16120deg_radius2_sig.eps}
        \caption{Image with sig.~pixels}
      \end{subfigure}
      \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSep16120deg_radius2_logp.eps}
        \caption{$-\log p$}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSep16120deg_radius2_nullMean.eps}
        \caption{Empirical null mean}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSep16120deg_radius2_nullStd.eps}
        \caption{Empirical null std}
      \end{subfigure}
    }
  \caption{Resulting inference when using the empirical null filter with radius $r=\addNumber{../figures/inference/DefectDetectSep16120deg_radius2.txt}$. a) Highlighted in red are significant pixels when controlling the FDR at the $2\sigma$ level. b) $p$-values on the log scale. c) and d) Empirical null mean and std.}
  \label{fig:inference_sep16120deg_inference_radius2}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSep16120deg_radius3_sig.eps}
        \caption{Image with sig.~pixels}
      \end{subfigure}
      \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSep16120deg_radius3_logp.eps}
        \caption{$-\log p$}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSep16120deg_radius3_nullMean.eps}
        \caption{Empirical null mean}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSep16120deg_radius3_nullStd.eps}
        \caption{Empirical null std}
      \end{subfigure}
    }
  \caption{Resulting inference when using the empirical null filter with radius $r=\addNumber{../figures/inference/DefectDetectSep16120deg_radius3.txt}$. a) Highlighted in red are significant pixels when controlling the FDR at the $2\sigma$ level. b) $p$-values on the log scale. c) and d) Empirical null mean and std.}
  \label{fig:inference_sep16120deg_inference_radius3}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSep16120deg_radius4_sig.eps}
        \caption{Image with sig.~pixels}
      \end{subfigure}
      \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSep16120deg_radius4_logp.eps}
        \caption{$-\log p$}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSep16120deg_radius4_nullMean.eps}
        \caption{Empirical null mean}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSep16120deg_radius4_nullStd.eps}
        \caption{Empirical null std}
      \end{subfigure}
    }
  \caption{Resulting inference when using the empirical null filter with radius $r=\addNumber{../figures/inference/DefectDetectSep16120deg_radius4.txt}$. a) Highlighted in red are significant pixels when controlling the FDR at the $2\sigma$ level. b) $p$-values on the log scale. c) and d) Empirical null mean and std.}
  \label{fig:inference_sep16120deg_inference_radius4}
\end{figure}

The empirical null filter was applied to the \texttt{Sep16 120deg} image as shown at the start of the chapter. Figure \ref{fig:inference_sep16120deg_zfiltered} shows the filtered $z$ images for various kernel radiuses. For $r=\addNumber{../figures/inference/DefectDetectSep16120deg_radius1.txt}$, the resulting $z$ image is smooth and treated everything as the null. When a large enough radius was used, the defects are highlighted by the higher $z$ statistics and everything else remain unchanged. Figures \ref{fig:inference_sep16120deg_inference_radius1} to \ref{fig:inference_sep16120deg_inference_radius4} shows the resulting inference along with the empirical null parameters. The figures shows that for large enough kernel radius, the filter irons out the false positives observed at the start of this chapter. All of the defects were successfully highlighted by the hypothesis test.

False positives were detected on the bottom right of the 3D printed sample. This is the result of the null distribution being bimodal where each face has a different distribution. Figure \ref{fig:inference_cornerSelect} illustrates this. The geometry of the faces and the kernel is such it results in one of the faces to be treated as non-null and then tested positive, falsely so.

\begin{figure}[htp]
  \centering
    \centerline{
      \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/cornerSelect.png}
        \caption{$z$ image with kernel of radius $r=120$}
      \end{subfigure}
      \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/cornerSelectHisto.png}
        \caption{Histogram of $z$ in the kernel}
      \end{subfigure}
    }
  \caption{A kernel of radius $r=120$ is placed over the bottom right of a unfiltered $z$ image. The histogram of the $z$ statistics captured by the kernel shows a bimodal distribution.}
  \label{fig:inference_cornerSelect}
\end{figure}

The false positives were tackled by manually segmenting the image further as shown in Figure \ref{fig:inference_segmentFurther}. It was segmented using the edges of the projection of the 3D printed object. The empirical null filter was then used on each segment independently, ignoring pixels outside the segment the filter is working on. The resulting filtered segments were stitched together to form the resulting filtered image. By using this method, the resulting inference is shown in Figure \ref{fig:inference_subroisep16120deg_inference_radius4}. It can be seen that the empirical null mean does change face to face, resulting in a clear boundary set by the edges. Most of false positives around the corners and edges have been eliminated.

\begin{figure}[htp]
  \centering
  \includegraphics[width=0.49\textwidth]{../figures/inference/segment.JPG}
  \caption{The $z$ was segmented further into 7 segments.}
  \label{fig:inference_segmentFurther}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSubRoiSep16120deg_radius4_sig.eps}
        \caption{Image with sig.~pixels}
      \end{subfigure}
      \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSubRoiSep16120deg_radius4_logp.eps}
        \caption{$-\log p$}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSubRoiSep16120deg_radius4_nullMean.eps}
        \caption{Empirical null mean}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/DefectDetectSubRoiSep16120deg_radius4_nullStd.eps}
        \caption{Empirical null std}
      \end{subfigure}
    }
  \caption{Resulting inference when using the empirical null filter, on each segment independently, with radius $r=\addNumber{../figures/inference/DefectDetectSubRoiSep16120deg_radius4.txt}$. a) Highlighted in red are significant pixels when controlling the FDR at the $2\sigma$ level. The $z$ statistics from all segments were combined in the BH procedure. b) $p$-values on the log scale. c) and d) Empirical null mean and std.}
  \label{fig:inference_subroisep16120deg_inference_radius4}
\end{figure}

\section{Conclusion}

Clusters of significant pixels should raise suspicions. One could create a binary image, assigning a Boolean value whether that pixel is tested positive or not. A binary image filter, such as erode followed by a dilate, can be used to remove isolate significant pixels and emphasise the cluster of significant pixels.