### 2019-08-05 reports/thesis/book.tex
compound poisson

\documentclass[12pt, a4paper]{memoir}
%\documentclass[12pt, a4paper, oldfontcommands]{memoir}
%\documentclass[12pt, a4paper]{report}

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{amsbsy}
\usepackage{graphicx}
\usepackage[outdir=./]{epstopdf}
\usepackage[round]{natbib} 
\usepackage{url}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{hyphenat}
\usepackage{pdfpages}
\usepackage{MnSymbol} %udots
\usepackage{afterpage} %\afterpage{\clearpage}
\usepackage{siunitx} %for measurements
\usepackage[figuresright]{rotating}
\usepackage{multirow} %for multirow tables
\usepackage{mathtools} %multlined
\usepackage{longtable}
\usepackage{bibentry}

\setcounter{tocdepth}{2}
\setcounter{secnumdepth}{2}

\let\footruleskip\undefined
\usepackage{fancyhdr}
\fancypagestyle{plain}{%
\fancyhf{}% clears all header and footer fields
\fancyhead[LE,RO]{\thepage}%
\renewcommand{\headrulewidth}{0pt}%
\renewcommand{\footrulewidth}{0pt}}

\DeclareMathOperator{\expfamily}{ExpFamily}
\DeclareMathOperator{\expectation}{\mathbb{E}}
\DeclareMathOperator{\variance}{\mathbb{V}ar}
\DeclareMathOperator{\cov}{\mathbb{C}ov}
\DeclareMathOperator{\corr}{\mathbb{C}orr}
\DeclareMathOperator{\bernoulli}{Bernoulli}
\DeclareMathOperator{\betaDist}{Beta}
\DeclareMathOperator{\dirichlet}{Dir}
\DeclareMathOperator{\bin}{Bin}
\DeclareMathOperator{\MN}{Multinomial}
\DeclareMathOperator{\prob}{\mathbb{P}}
\DeclareMathOperator{\trace}{Tr}
\DeclareMathOperator{\normal}{N}
\DeclareMathOperator{\gammaDist}{Gamma}
\DeclareMathOperator{\poisson}{Poisson}
\DeclareMathOperator{\CPoisson}{CP\Gamma}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\RSS}{\mathrm{RSS}}
\newcommand{\euler}{\mathrm{e}}
\newcommand{\diff}{\mathrm{d}}
\newcommand{\T}{^\textup{T}}
\newcommand{\dotdotdot}{_{\phantom{.}\cdots}}
\newcommand{\BIC}{\mathrm{BIC}}
\newcommand{\AIC}{\mathrm{AIC}}

\newcommand{\subSize}{0.49\textwidth}
\newcommand{\mainSize}{0.8\textwidth}

\newcommand{\vect}[1]{\mathbf{#1}}
\newcommand{\vectGreek}[1]{\boldsymbol{#1}}
\newcommand{\matr}[1]{\mathsf{#1}}

\newcommand{\addNumber}[1]{\protect\input{#1}\unskip}
\newcommand{\inputNumber}[1]{\protect\input{#1}\unskip}

\DeclareSIUnit\pixel{px}
\DeclareSIUnit\adu{ADU}

\begin{document}\sloppy

\begin{titlingpage}
\centering
{\LARGE Characterisation of Computed Tomography Noise in Projection Space with Applications to Additive Manufacturing \par}
\vspace{1cm}
{\Large Sherman E.~Lo\par}
{\Large Doctor of Philosophy in Statistics\par}
\vfill
\includegraphics[width = 0.75\textwidth]{../figures/frontCover.jpg}
\vfill
{\Large University of Warwick\par}
{\Large Department of Statistics\par}
{\Large September 2019\par}
\end{titlingpage}

\frontmatter

\newpage
\tableofcontents*
\newpage
\listoffigures
\newpage
\listoftables

\chapter{Acknowledgements}
I would like to thank my supervisors Julia Brettschneider and Tom Nichols. I would also like to thank my calibrators, as part of the Inside-Out group, in Warwick Statistics, Clair Barnes, Wilfred Kendall and Audrey Kueh, and the Warwick Manufacturing Group, Greg Gibbons, Jay Warnett and Mark Williams.

This work is funded by the EPSRC and MRC Centre for Doctoral Training in Next Generation Statistical Science: The Oxford-Warwick Statistics Programme (EP/L016710/1). I would like to thank my cohort for the friendship and collaboration during the programme: Nathan Cunningham, Giuseppe di Benedetto, Beniamino Hadj-Amar, Jack Jewson, Ella Kaye, Leon Law, Kaspar Martens, Marcin Mider, Xenia Miscouridou, Paul Vanetti and Andi Wang.

\chapter{Declaration}
This thesis is submitted to the University of Warwick in support of my application for the degree of Doctor of Philosophy. It has been composed by myself and has not been submitted in any previous application for any degree.

The work presented (including data generated and data analysis) was carried out by the author except in the cases outlined below:
\begin{itemize}
  \item The fabrication, x-ray computed tomography acquisition and simulation of test samples were done by Greg Gibbons and Jay Warnett, from the Warwick Manufacturing Group, as part of Inside-out: Statistical methods for Computed Tomography validation of complex structures in Additive Layer Manufacturing funded by EPSRC (EP/K031066/1).
\end{itemize}

Parts of this thesis have been published by the author:
\begin{itemize}
  \item\nobibliography*\bibentry{lo2019detection}
\end{itemize}

\chapter{Abstract}
X-ray computed tomography (CT) can be used for defect detection in 3D printing. The object is scanned at multiple angles to reconstruct the object in 3D space. The process can be time consuming. The aim of this project was to investigate if it is possible to conduct defect detection in projection space from a single scan to speed up the defect detection procedure. To do this, a 3D printed sample was manufactured with voids to see if they can be detected.

X-ray photons behave randomly. Hence to do defect detection pixel by pixel, uncertainty must be taken into account. A compound Poisson model was used to model the grey valyes in a pixel. It assumed that photon arrivals are a Poisson process with Gamma distributed energy. This resulted in a linear relationship between the mean and variance of the grey value, which was used for variance prediction and to quantify the uncertainty.

Software, called \texttt{aRTist}, was used to simulate the scan and it was compared with the x-ray acquisition under the face of uncertainty. However the software, and the information provided to it, was not perfect which led to model misspecification and incorrect inference. The empirical null filter was proposed. It adjust each pixel statistic so that inference was done by comparing each pixel with the majority of its local pixels.

\newpage

\mainmatter

\chapter{Introduction}

\chapter{Literature Review}
\input{chapters/1_literatureReview.tex}

\chapter{Data Collection}
\input{chapters/2_dataCollection.tex}

\chapter{Compound Poisson}
\input{chapters/3_compoundPoisson.tex}

\chapter{Variance Prediction}
\input{chapters/4_variancePrediction.tex}

\chapter{Inference}
\input{chapters/5_inference.tex}

\bibliographystyle{apalike}
\bibliography{../bib}

\begin{appendices}
\input{chapters/appendix.tex}
\end{appendices}

\end{document}